---
title: "Retrieve_format_barcode"
author: "D Stephens"
date: "2025-07-17"
output: html_document
---

# Setup Env and Libraries
```{r}

# Set working directory
setwd("C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis")

# Load Libraries
library(Biostrings)
library(taxize)
library(VennDiagram)
library(ggplot2)
library(sf)
library(purrr)
library(xml2)
library(rentrez)
library(gbfetch)
library(grid)
library(ggpattern)
library(scales)
library(ggpubr)
library(ggrepel)

library(dplyr)
library(tidyr)
library(stringr)
library(readxl)
library(selenider)
library(rvest)
library(httr)
library(jsonlite)

```

# Prepare Name Change Data
Takes British Lichen Society's Taxon Dictionary and places synonyms/historic names in an old name column and current name in the new name column. This will be used to homogenize sequence records once retrieved.
```{r}
# British Lichen Society Taxon Dictionary
bls <- read.csv("Updater/March_2026/Input_data/bls-taxon-dictionary-2026-03-03.csv")

# Join to DB to get old and new names
names <- bls %>% select(Name, Current.Taxon.Name, Synonyms)

# replace semicolons with commas
names <- names %>%
  mutate(Synonyms = gsub(";", ",", Synonyms))

# Split the taxon_list column into multiple rows based on the comma
names <- names %>%
  separate_rows(Synonyms, sep = ",")

# Trim any extra spaces
names <- names %>%
  mutate(Synonyms = trimws(Synonyms))

names <- names %>%
  # Gather Name and Synonyms into the old_name column
  pivot_longer(cols = c("Name", "Synonyms"), 
               names_to = "type", 
               values_to = "old_name") %>%
  filter(old_name != "") %>%  # Remove empty rows after pivoting
  select(old_name, new_name = Current.Taxon.Name)

# Split the old and new name values by spaces and extract the first two words
names$old_name <- sapply(strsplit(names$old_name, "\\s+"), function(words) {
  paste(words[1:min(2, length(words))], collapse = " ")
})

names$new_name <- sapply(strsplit(names$new_name, "\\s+"), function(words) {
  paste(words[1:min(2, length(words))], collapse = " ")
})

# odd characters
names$old_name <- gsub("\\?", "", names$old_name)
names$old_name <- gsub("[‘’'\"]", "", names$old_name )
names$new_name <- gsub("\\?", "", names$new_name)
names$new_name <- gsub("[‘’'\"]", "", names$new_name)
names <- names %>%
  mutate(old_name = trimws(old_name),
         new_name = trimws(new_name))

# Remove NA values
names$new_name <- gsub("NA", NA, names$new_name)
names <- na.omit(names)

# remove sp. and one word values in either old or new name
names$old_name <- gsub("\\s+sp\\.$", "", names$old_name)
names$new_name <- gsub("\\s+sp\\.$", "", names$new_name)

# Filter out entries that are only one word
names <- names[sapply(strsplit(names$old_name, "\\s+"), length) > 1, ]
names <- names[sapply(strsplit(names$new_name, "\\s+"), length) > 1, ]

names <- unique(names)

names$new_name <- gsub("Xanthmendoza fulva", "Xanthomendoza fulva", names$new_name)

# Write out names
write.csv(names, "Updater/March_2026/Output_data/name_changes.csv", row.names = F)

length(unique(bls$Current.Taxon.Name))

```

2751 current names

# Name Change Function
```{r}
fix_names <- function(sequence_df, names_df) {
  
  # 1) Summarize names_df so each old_name has a list of new_names
  map_summarized <- names_df %>%
    group_by(old_name) %>%
    summarize(new_names = list(unique(new_name)), .groups = "drop")
  
  # 2) Join onto sequence_df
  df_joined <- sequence_df %>%
    left_join(map_summarized, by = c("species" = "old_name"))
  # This yields a 'new_names' list-column for each row in sequence_df.
  
  # 3) Define your custom logic for one row
  fix_species_one_row <- function(species_val, seq_val, possible_new) {
    # CASE A: No match => keep original
    if (is.null(possible_new) || all(is.na(possible_new))) {
      return(tibble(species = species_val, sequence = seq_val))
    }
    
    # CASE B: Exactly 1 new name
    if (length(possible_new) == 1) {
      if (possible_new == species_val) {
        # same as old => no change
        return(tibble(species = species_val, sequence = seq_val))
      } else {
        # single new name => replace
        return(tibble(species = possible_new, sequence = seq_val))
      }
    }
    
    # CASE C: Multiple new names
    # If all new names == old => no change
    if (all(possible_new == species_val)) {
      return(tibble(species = species_val, sequence = seq_val))
    }
    
    # There's at least one new name different from the old:
    if (species_val %in% possible_new) {
      # Mixture of old + new => keep old
      return(tibble(species = species_val, sequence = seq_val))
    } else {
      # True split => remove this row entirely
      return(tibble())  # returns zero rows
    }
  }
  
  # 4) Apply fix_species_one_row to each row
  df_fixed <- df_joined %>%
    mutate(
      fixed = pmap(
        list(species, sequence, new_names),
        ~ fix_species_one_row(..1, ..2, ..3)
      )
    ) %>%
    # Remove the original columns we no longer need:
    select(-species, -sequence, -new_names) %>%
    # Expand the list-column
    unnest(cols = fixed)
  
  # 5) Return the final cleaned dataframe
  return(df_fixed)
}
```

# Sequence Data Retrieval
## GenBank
*Skip normally*
Check for which markers have the most species coverage
```{r}

# Enter key here
ncbi_api_key = "3250605898efe21223d0a656ad86a363c608"
Sys.setenv(ENTREZ_KEY = ncbi_api_key)

# Define your list of lichen species
lichen_species <- c("Cladonia rangiferina", "Usnea barbata", "Parmelia sulcata")
lichen_species <- sort(unique(names$old_name, names$new_name))

# Initialize an empty dataframe to store metadata
all_metadata <- data.frame()
i = 1000
  
# Loop through each lichen species and fetch metadata
for (species in lichen_species[1000:length(lichen_species)]) {
  cat("Fetching metadata for:", species, "\n")
  cat("Species: ", i, " of ", length(lichen_species), "\n")
  
  # Construct the query for GenBank
  query <- paste0(species, "[ORGN]")  # Search for the species in the organism field
  
  # Fetch metadata
  species_metadata <- tryCatch(
    {
      fetch_metadata(query = query, chunk_size = 10, max_tries = 10, verbose = TRUE)
    },
    error = function(e) {
      cat("Error fetching metadata for", species, ":", e$message, "\n")
      return(NULL)
    }
  )
  
  i = i + 1
  
  # Append metadata to the main dataframe
  if (!is.null(species_metadata)) {
    
    all_metadata <- bind_rows(all_metadata, species_metadata)
  }

}

all_metadata <- all_metadata %>% select("gi", "accession", "taxid", "title", "slen", "subtype", "subname", "species"  )

# write.csv(all_metadata, "Updater/March_2026/Output_data/genbank_meta_full.csv")

# Country filtering
euro_countries <- c("Russia", 'Germany', "United Kingdom", "France", "Italy", "Spain", "Poland", "Ukraine", "Romania", "Netherlands", "Belgium", "Czech Republic", "Sweden", "Portugal", "Greece", "Hungary", "Austria", "Belarus", "Switzerland", "Bulgaria", "Serbia", "Denmark", "Finland", "Norway", "Slovakia", "Ireland", "Croatia", "Bosnia and Herzegovina", "Moldova", "Lithuania", "Albania", "Slovenia", "Latvia", "North Macedonia", "Estonia", "Luxembourg", "Montenegro", "Malta", "Iceland", "Andorra", "Liechtenstein", "Monaco", "San Marino", "Holy See")

# Extract country information into a new column
all_metadata$country <- mapply(function(subtype, subname) {
  # Split the strings
  subtype_fields <- unlist(strsplit(subtype, "\\|"))
  subname_values <- unlist(strsplit(subname, "\\|"))
  
  # Find the position of 'country' in subtype
  country_index <- which(subtype_fields == "country")
  
  # Return country value if it exists
  if (length(country_index) == 1 && length(subname_values) >= country_index) {
    return(subname_values[country_index])
  } else {
    return(NA)
  }
}, all_metadata$subtype, all_metadata$subname)

all_metadata$country <- str_split_i(all_metadata$country, ":", 1) 

# Filter Countries
all_metadata_euro <- all_metadata %>%
  filter(country %in% euro_countries)

# Genetic Marker Filter
marker_mapping <- list(
  "ITS" = c("ITS", "internal transcribed spacer", "intergenic spacer region"),
  "ITS1" = c("ITS1", "internal transcribed spacer 1"),
  "ITS2" = c("ITS2", "internal transcribed spacer 2"),
  "5.8S" = c("5.8S", "5.8S ribosomal RNA gene"),
  "COX1" = c("cox1", "COX1", "yclooxygenase-1", "ytochrome c oxidase subunit 1"),
  "EF1a" = c("EF1a", "elongation factor 1a-like gene", "EF-1", "elongation factor 1"),
  "TEF1" = c("TEF1", "tef1", "Tef1", "tef-1"),
  "Sog2" = c("Sog2"),
  "28S-18S" = c("28S-18S", "28S-18S ribosomal RNA intergenic spacer, partial sequence"),
  "IGS" = c("IGS", "28S-18S ribosomal RNA intergenic spacer", "intergenic spacer"),
  "GPD" = c("GPD", "gpd"),
  "pks2" = c("pks2"),
  "GAPDH" = c("GAPDH", "glyceraldehyde-3-phosphate dehydrogenase"),
  "TSR1" = c("TSR1"),
  "GZU" = c("GZU"),
  "RPS3aE" = c("RPS3aE"),
  "benA" = c("benA"),
  "hypothetical protein" = c("hypothetical protein"),
  "MCM7" = c("MCM7", "mcm7", "Mcm7", "minichromosome maintenance complex component 7-like gene"),
  "TUB2" = c("beta-tubulin gene", "beta tubulin", "beta-tubulin-like gene", "beta-tubilin", "beta-tubulin", "TUB2", "beta-tublin"),
  "mRNA sequence" = c("mRNA sequence"),
  "nucLSU" = c("large subunit ribosomal RNA", "28S"),
  "nucSSU" = c("18S"),
  "mtSSU" = c("small subunit ribosomal RNA gene, partial sequence; mitochondrial"),
  "40S" = c("40S"),
  "23S" = c("23S"),
  "12S" = c("12S"),
  "16S" = c("16S"),
  "60S" = c("60S"),
  "RPB2" = c("RPB2", "DNA-directed RNA polymerase II", "RNA polymerase II second largest subunit", "RNA polymerase 2 largest subunit gene"),
  "RPB1" = c("RPB1", "RNA polymerase II largest subunit", "RNA polymerase II large subunit", "RNA polymerase II large subunit"),
  "mitochondrion" = c("mitochondrion"),
  "whole genome" = c("whole genome"),
  "LNS2" = c("LNS2"),
  "PGK" = c("PGK", "3-phosphate dehydrogenase", "3 phosphate dehydrogenas", "pgk", "phosphoglycerate kinase 1"),
  "COX2" = c("COX2", "cox2", "yclooxygenase-2", "yclooxygenase-2"),
  "PKS" = c("PKS", "pks"),
  "ks1" = c("ks1"),
  "25S" = c("25S"),
  "elongation factor 3" = c("elongation factor 3"),
  "chromosome" = c("chromosome"),
  "mep1A" = c("mep1A"),
  "ChsG" = c("ChsG"),
  "heat shock protein" = c("heat shock protein"),
  "hsp72" = c("hsp72"),
  "AnG22" = c("AnG22"),
  "AnG25" = c("AnG25"),
  "AnG23" = c("AnG23"),
  "mep1B" = c("mep1B"),
  "mep2" = c("mep2"),
  "actin" = c("actin"),
  "elongation factor 2" = c("elongation factor 2"),
  "26S rRNA" = c("26S rRNA"),
  "TOP1" = c("TOP1", "DNA topoisomerase I", "DNA topoisomerase 1")
)

# Function to add marker columns and consolidate them
add_marker_columns <- function(data, marker_mapping) {
  # Create an empty dataframe for marker columns
  marker_columns <- data.frame(matrix(0, nrow = nrow(data), ncol = length(marker_mapping)))
  colnames(marker_columns) <- names(marker_mapping)
  
  # Loop through each marker in the mapping
  for (marker in names(marker_mapping)) {
    # Create a regex pattern for all synonyms of the marker
    pattern <- paste(marker_mapping[[marker]], collapse = "|")
    
    # Add a column for the marker
    marker_columns[[marker]] <- as.integer(grepl(pattern, data$title, ignore.case = TRUE))
  }
  
  # Bind marker columns to the original dataframe
  data <- cbind(data, marker_columns)
  return(data)
}

# Apply the function to add marker columns
all_metadata_euro <- add_marker_columns(all_metadata_euro, marker_mapping)

# Identify rows where no markers matched
unmatched_rows <- all_metadata_euro %>%
  filter(rowSums(across(names(marker_mapping))) == 0)

# 1433 entries without gene ID,in 52,500 records

# Create Marker_Combination column explicitly
all_metadata_euro <- all_metadata_euro %>%
  mutate(Marker_Combination = apply(across(names(marker_mapping)), 1, function(row) {
    if (sum(row) > 1) {
      paste(names(row)[which(row == 1)], collapse = ", ")
    } else {
      NA
    }
  }))

write.csv(all_metadata_euro, "Updater/March_2026/Output_data/genbank_meta_euro_marker.csv")


# Summarize by Marker_Combination
combination_summary <- all_metadata_euro %>%
  filter(!is.na(Marker_Combination)) %>%  # Only rows with non-NA Marker_Combination
  group_by(Marker_Combination) %>%  # Group by Marker_Combination
  summarize(
    Species_Count = n_distinct(species),
    .groups = "drop"
  ) %>%
  dplyr::rename(Marker = Marker_Combination)  # Rename column for clarity

# Group by species and calculate unique species count per marker
species_marker_summary <- all_metadata_euro %>%
  group_by(species) %>%
  summarize(across(names(marker_mapping), max)) %>% # Use max to detect presence across rows for a species
  summarize(across(names(marker_mapping), sum)) %>% # Sum the presence across species
  pivot_longer(cols = everything(), names_to = "Marker", values_to = "Species_Count") %>%
  arrange(desc(Species_Count))

# Bind individual and combination summaries
combined_and_individual <- bind_rows(species_marker_summary, combination_summary)

# Get the top 20 markers
top_20_markers <- combined_and_individual[order(-combined_and_individual$Species_Count), ][1:20, ]

```

### Plot GenBank Prelim Data
```{r}

ggplot(top_20_markers, aes(x = reorder(Marker, -Species_Count), y = Species_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Top 20 Marker Combinations and Individual Markers by Species Count",
    x = "Marker",
    y = "Species Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Filter Markers
targeted_markers <- c("COX1", "COX2", "TEF1","ITS2", "ITS1", "IGS", "nucLSU", "nucSSU", "mtSSU", "mtLSU","IGS", "TUB2", "RPB1", "RPB2", "TEF1", "PGK")

targeted_markers_df <- combined_and_individual %>% filter(Marker %in% targeted_markers)

# Plot the top 20 markers
ggplot(targeted_markers_df, aes(x = reorder(Marker, -Species_Count), y = Species_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "",
    x = "Marker",
    y = "Number of Species"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# ITS
## UNITE
Two methods to download data, can compile both of these
### Method 1
#### Species Webscrape
*done*
Searches on UNTIE sequence tab for species then grabs records by text
```{r}
# Start a new session
options(httr2_timeout = 500) 
session <- selenider_session("selenium", browser = "chrome")

UNITE_sequences <- data.frame()

datasets <- c("unite", "insd")

base_url_for <- function(dataset) {
  paste0(
    "https://unite.ut.ee/search.php?qresult_seq=yes",
    "&dataset=", dataset,
    "&taxon_name_seq=&taxon_id_seq=&taxon_name_repr_seq=&accno_seq=",
    "&it_seq=&it_id_seq=&it_repr_seq=&country_seq=&country_id_seq=&country_repr_seq=",
    "&insd_organism=&source_seq=&ecm_lineage_seq=&region_seq=1#fndtn-panel2"
  )
}

ENTER      <- "\uE007"
ARROW_DOWN <- "\uE015"

# Select FIRST suggestion (normal path): type + ENTER only
select_first_dropdown_option <- function(species, sleep_after_type = 0.6) {
  inp <- s("#taxon_name_seq")
  try(inp %>% elem_click(), silent = TRUE)
  Sys.sleep(1)

  inp %>% elem_clear_value() %>% elem_send_keys(species)
  Sys.sleep(sleep_after_type)

  # NORMAL PATH: no arrows, just enter
  inp %>% elem_send_keys(ENTER)
  Sys.sleep(1)
}

# Select Nth suggestion (N>=2): type + ArrowDown^(N-1) + ENTER
select_nth_dropdown_option <- function(species, n, sleep_after_type = 0.6, sleep_between = 0.08) {
  stopifnot(n >= 2)

  inp <- s("#taxon_name_seq")
  try(inp %>% elem_scroll_to(), silent = TRUE)
  try(inp %>% elem_click(), silent = TRUE)
  Sys.sleep(1)

  inp %>% elem_clear_value() %>% elem_send_keys(species)
  Sys.sleep(sleep_after_type)

  for (k in seq_len(n - 1L)) {
    inp %>% elem_send_keys(ARROW_DOWN)
    Sys.sleep(sleep_between)
  }

  inp %>% elem_send_keys(ENTER)
  Sys.sleep(1)
}

# Simple subtaxa click (no ensure; avoids toggling on/off loops)
select_subtaxa <- function() {
  try({
    cb <- s("#parent_taxon_seq")
    chk <- try(cb %>% elem_prop("checked"), silent = TRUE)
    is_checked <- isTRUE(chk) || identical(chk, "true") || identical(chk, "checked")
    if (!is_checked) cb %>% elem_click()
  }, silent = TRUE)
}

# Read the selected taxon label from the chip (this is the UNITE taxon name selection)
read_selected_unite_label <- function(timeout = 2.5, poll = 0.1) {
  t0 <- Sys.time()
  css <- "#taxon_name_repr_div_seq > span:nth-child(1)"
  repeat {
    lab <- try(s(css) %>% elem_text(), silent = TRUE)
    if (!inherits(lab, "try-error")) {
      lab <- trimws(as.character(lab))
      if (nzchar(lab)) return(lab)
    }
    if (as.numeric(difftime(Sys.time(), t0, units = "secs")) > timeout) break
    Sys.sleep(poll)
  }
  NA_character_
}

# Base name = first 2 tokens (Genus species), ignoring subsp./var./f./ssp. etc.
base_binomial <- function(x) {
  x <- norm_taxon(x)

  # strip common infraspecific markers and everything after them
  x <- gsub("\\b(subsp\\.|ssp\\.|var\\.|f\\.|forma)\\b.*$", "", x)

  # keep only first two words (genus + species epithet)
  toks <- unlist(strsplit(x, "\\s+"))
  if (length(toks) >= 2) paste(toks[1], toks[2]) else x
}

read_first_row_species <- function(timeout = 2.5, poll = 0.1) {
  t0 <- Sys.time()
  css <- "#panel2_results > div > div > table > tbody > tr:nth-child(1) > td:nth-child(3)"
  repeat {
    tx <- try(s(css) %>% elem_text(), silent = TRUE)
    if (!inherits(tx, "try-error")) {
      tx <- trimws(as.character(tx))
      if (nzchar(tx)) return(tx)
    }
    if (as.numeric(difftime(Sys.time(), t0, units = "secs")) > timeout) break
    Sys.sleep(poll)
  }
  NA_character_
}

# Normalize taxon strings so authorship/quotes don’t break matching
norm_taxon <- function(x) {
  x <- tolower(trimws(x))
  x <- gsub("[‘’'\"]", "", x)         # remove curly/straight quotes
  x <- gsub("\\s+", " ", x)           # collapse whitespace
  x
}

choose_dropdown_option <- function(input_css,
                                   option_index = 1L,
                                   type_text = NULL,
                                   sleep_after_type = 1,
                                   sleep_between = 1) {
  inp <- s(input_css)

  try(inp %>% elem_scroll_to(), silent = TRUE)
  try(inp %>% elem_click(), silent = TRUE)  # focus
  Sys.sleep(1)

  if (!is.null(type_text)) {
    inp %>% elem_clear_value() %>% elem_send_keys(type_text)
    Sys.sleep(sleep_after_type)
  }

  # ALWAYS trigger/open the dropdown suggestions
  # (On UNITE, ENTER alone often doesn't open/select a suggestion)
  inp %>% elem_send_keys(ARROW_DOWN)
  Sys.sleep(1)

  # Now choose the option_index-th suggestion:
  # option_index == 1 -> already on first suggestion, just ENTER
  # option_index > 1  -> move down (option_index - 1) more times, then ENTER
  if (option_index > 1L) {
    for (j in seq_len(option_index - 1L)) {
      inp %>% elem_send_keys(ARROW_DOWN)
      Sys.sleep(sleep_between)
    }
  }

  inp %>% elem_send_keys(ENTER)
  Sys.sleep(1)

  invisible(TRUE)
}

read_total_records <- function(timeout = 2.5, poll = 0.1) {
  t0 <- Sys.time()
  css <- "#panel2_results > div > div > div > div:nth-child(1) > span > b"
  repeat {
    tx <- try(s(css) %>% elem_text(), silent = TRUE)
    if (!inherits(tx, "try-error")) {
      tx <- trimws(as.character(tx))
      if (nzchar(tx)) {
        n <- suppressWarnings(as.integer(gsub("\\D+", "", tx)))
        if (!is.na(n)) return(n)
      }
    }
    if (as.numeric(difftime(Sys.time(), t0, units = "secs")) > timeout) break
    Sys.sleep(poll)
  }
  NA_integer_
}

click_results_page <- function(page_num) {
  # page 2 -> li:nth-child(2), page 3 -> li:nth-child(3), etc.
  css <- paste0(
    "#panel2_results > div > div > div > div:nth-child(2) > div > ul > li:nth-child(",
    page_num,
    ") > a"
  )
  ok <- try(s(css) %>% elem_click(), silent = TRUE)
  if (inherits(ok, "try-error")) FALSE else TRUE
}

scrape_current_results_page <- function(species, dataset, selected_label) {

  species_info <- tryCatch({
    s("#panel2_results > div > div > table > tbody") %>% elem_text()
  }, error = function(e) NA_character_)

  if (is.na(species_info) || !nzchar(species_info)) {
    return(NULL)
  }

  beginning <- "Accession number UNITE taxon name Country Source"
  end <- "TWEETS"

  clipped_text <- sub(paste0(".*", beginning), "", species_info)
  clipped_text <- sub(paste0(end, ".*"), "", clipped_text)

  rows <- unlist(strsplit(clipped_text, "\n"))
  rows <- rows[rows != ""]
  rows <- trimws(rows)

  n_records <- length(rows)

  if (!n_records) {
    return(NULL)
  }

  rows_clean <- sub("^\\d+\\.\\s*", "", rows)
  split_rows <- strsplit(rows_clean, "\\s+")

  accession_numbers <- vapply(split_rows, function(x) x[1], character(1))

  countries_sources <- vapply(
    split_rows,
    function(x) paste(x[4:length(x)], collapse = " "),
    character(1)
  )

  countries <- str_split_i(countries_sources, "Fungus:", 1) %>% trimws()
  sources   <- str_split_i(countries_sources, "Fungus:", 2) %>% trimws()

  df <- data.frame(
    Accession_Number = accession_numbers,
    Taxon_Name       = species,
    Country          = countries,
    Source           = sources,
    sequence         = NA_character_,
    Marker           = NA_character_,
    UNITE_taxon_name = selected_label,
    database         = toupper(dataset),
    Lineage          = NA_character_,
    Specimen_Voucher = NA_character_,
    Sequence_Date    = NA_character_,
    stringsAsFactors = FALSE
  )

  Sys.sleep(1)

  for (i in seq_len(n_records)) {

    selector <- if (i == 1) {
      "#panel2_results > div > div > table > tbody > tr > td:nth-child(2) > a"
    } else {
      paste0(
        "#panel2_results > div > div > table > tbody > tr:nth-child(",
        i,
        ") > td:nth-child(2) > a"
      )
    }

    record_temp <- s(selector)

    click_ok <- try(record_temp %>% elem_click(), silent = TRUE)

    if (inherits(click_ok, "try-error")) {
      next
    }

    all_windows <- session$driver$window_handles()

    if (length(all_windows) < 2L) {
      next
    }

    session$driver$switch_to_window(all_windows[[2]][1])
    
        # Give browser a moment to switch windows
    Sys.sleep(1.5)
    
    text <- wait_for_text(
      "#main > div:nth-child(3) > div > div > div > table:nth-child(2) > tbody",
      timeout = 25,
      poll = 0.5
    )
    
    if (is.na(text)) {
      message(sprintf("[%s] %s: details page table never loaded (row %d/%d)",
                      toupper(dataset), species, i, n_records))
    
      session$driver$close_window()
      session$driver$switch_to_window(all_windows[[1]][1])
      Sys.sleep(0.75)
      next
    }

    df[i, "Marker"] <- str_match(
      text,
      "Sequenced regions\\s*(.*)\\n"
    )[, 2]

    seq_line <- str_match(
      text,
      "Sequence \\s*(.*)\\n"
    )[, 2]

    df[i, "sequence"] <- str_match(
      seq_line,
      "\\)\\s*(.*)"
    )[, 2]

    session$driver$close_window()

    session$driver$switch_to_window(all_windows[[1]][1])

    Sys.sleep(0.75)
  }

  df
}

retry_silent <- function(expr, tries = 3, sleep = 1) {
  for (k in seq_len(tries)) {
    out <- try(eval.parent(substitute(expr)), silent = TRUE)
    if (!inherits(out, "try-error")) return(out)
    Sys.sleep(sleep)
  }
  out
}

make_absolute_url <- function(href, base = "https://unite.ut.ee/") {
  href <- trimws(as.character(href))
  if (!nzchar(href)) return(NA_character_)

  # skip javascript links
  if (grepl("^javascript:", href, ignore.case = TRUE)) return(NA_character_)

  # already absolute
  if (grepl("^https?://", href, ignore.case = TRUE)) return(href)

  # protocol-relative
  if (grepl("^//", href)) return(paste0("https:", href))

  # relative path
  paste0(base, sub("^/+", "", href))
}

is_bad_href <- function(x) {
  is.na(x) || !nzchar(x)
}

wait_for_text <- function(css, timeout = 20, poll = 0.5) {
  start <- Sys.time()

  repeat {
    txt <- try(s(css) %>% elem_text(), silent = TRUE)

    if (!inherits(txt, "try-error")) {
      txt <- trimws(as.character(txt))
      if (nzchar(txt)) return(txt)
    }

    if (as.numeric(difftime(Sys.time(), start, units = "secs")) > timeout) {
      return(NA_character_)
    }

    Sys.sleep(poll)
  }
}

# species list
species_list <- sort(unique(c(names$old_name, names$new_name)))

max_variants <- 15
n = 6261 
species_fail_count <- 0
skip_species <- FALSE

for (species in species_list[6261:length(species_list)]) {

  print(paste0("Species ", n, " of ", length(species_list)))

  species_fail_count <- 0
  skip_species <- FALSE

  for (dataset in datasets) {

    first_label <- NA_character_
    last_label  <- NA_character_

    all_variant_results <- list()

    # before the variant loop (per species + dataset)
    seen_labels <- character(0)

    for (variant_i in seq_len(max_variants)) {

      # RESET state each variant so the taxon chip isn't "stuck"
      open_url(base_url_for(dataset))

      s("#region_seq_all") %>% elem_click()
      select_subtaxa()
      Sys.sleep(0.2)

      go_button <- s("#seq_search_full > div > div > div:nth-child(4) > div:nth-child(5) > input")

      # --- Select dropdown variant ---
      if (variant_i == 1L) {
        # normal path: ENTER only
        select_first_dropdown_option(species, sleep_after_type = 0.6)
      } else {
        # variant 2 => 1 arrow; variant 3 => 2 arrows; etc.
        select_nth_dropdown_option(
          species,
          n = variant_i,
          sleep_after_type = 0.6,
          sleep_between = 0.1
        )
      }

      # READ CHIP LABEL HERE
      lab <- read_selected_unite_label()

      if (is.na(lab) || !nzchar(lab)) {
        species_fail_count <- species_fail_count + 1

        message(sprintf(
          "[%s] %s: could not read chip label at variant %d (species fail %d/2)",
          toupper(dataset), species, variant_i, species_fail_count
        ))

        if (species_fail_count >= 2) {
          message(sprintf(
            "%s: chip label failed twice across datasets/variants -> skipping species",
            species
          ))
          skip_species <- TRUE
          break
        }

        next
      }

      # reset after a successful chip read
      species_fail_count <- 0

      lab_norm <- norm_taxon(lab)

      if (lab_norm %in% seen_labels) {
        message(sprintf(
          "[%s] %s: dropdown label '%s' already seen at variant %d -> stopping search",
          toupper(dataset), species, lab, variant_i
        ))
        break
      }

      seen_labels <- c(seen_labels, lab_norm)

      # If label didn't change from last time, don't do work; just continue scanning variants
      if (!is.na(last_label) && identical(norm_taxon(lab), norm_taxon(last_label))) {
        message(sprintf(
          "[%s] %s: label unchanged ('%s') at variant %d -> continuing",
          toupper(dataset), species, lab, variant_i
        ))
        next
      }

      last_label <- lab

      # --- Decide whether to collect this variant ---
      # Condition: base name matches query (ignoring subsp/var/f/etc.)
      if (base_binomial(lab) != base_binomial(species)) {
        message(sprintf(
          "[%s] %s: variant %d label '%s' base '%s' != query base '%s' -> not collecting, continuing",
          toupper(dataset), species, variant_i, lab,
          base_binomial(lab), base_binomial(species)
        ))
        next
      }

      # Run GO and check records
      Sys.sleep(0.5)
      go_button %>% elem_click()
      Sys.sleep(0.5)

      panel_txt <- try(s("#panel2_results") %>% elem_text(), silent = TRUE)

      if (!inherits(panel_txt, "try-error") &&
          grepl("No records found", panel_txt, ignore.case = TRUE)) {
        message(sprintf(
          "[%s] %s: variant %d ('%s') -> no records, continuing",
          toupper(dataset), species, variant_i, lab
        ))
        next
      }

      total_hits <- read_total_records()

      # If we can't read it, assume just 1 page (safe fallback)
      if (is.na(total_hits)) total_hits <- 50L

      n_pages <- ceiling(total_hits / 50)

      page_dfs <- list()

      # Page 1 (already loaded)
      df_p1 <- scrape_current_results_page(species, dataset, lab)
      if (!is.null(df_p1) && nrow(df_p1) > 0) {
        page_dfs[[length(page_dfs) + 1L]] <- df_p1
      }

      # Pages 2..n_pages
      if (n_pages > 1) {
        for (p in 2:n_pages) {

          ok <- click_results_page(p)
          Sys.sleep(1)

          if (!ok) {
            message(sprintf(
              "[%s] %s: couldn't click page %d/%d; stopping pagination for this variant",
              toupper(dataset), species, p, n_pages
            ))
            break
          }

          df_p <- scrape_current_results_page(species, dataset, lab)

          if (!is.null(df_p) && nrow(df_p) > 0) {
            page_dfs[[length(page_dfs) + 1L]] <- df_p
          } else {
            message(sprintf(
              "[%s] %s: page %d returned no parsable rows; continuing",
              toupper(dataset), species, p
            ))
          }
        }
      }

      df_variant <- if (length(page_dfs)) bind_rows(page_dfs) else NULL

      if (!is.null(df_variant) && nrow(df_variant) > 0) {
        all_variant_results[[length(all_variant_results) + 1L]] <- df_variant
        message(sprintf(
          "[%s] %s: collected %d records for variant %d ('%s')",
          toupper(dataset), species, nrow(df_variant), variant_i, lab
        ))
      } else {
        message(sprintf(
          "[%s] %s: variant %d ('%s') produced no parsable records; continuing",
          toupper(dataset), species, variant_i, lab
        ))
      }
    } # end variant loop

    if (length(all_variant_results)) {
      UNITE_sequences <- bind_rows(UNITE_sequences, bind_rows(all_variant_results))
    }

    if (skip_species) {
      break
    }
  }

  n <- n + 1

  if (skip_species) {
    next
  }
}


# Wriet out file
# write.csv(UNITE_sequences, "Updater/March_2026/Output_data/UNITE_sequences_all_LAT.csv")

```

Cerothallia luteoalba

index - 1344

n Started at 885 instead of 880 so subtract 5

Buellia atrataOrphniospora

#### Clean webscrape
```{r}

# Read in file
UNITE_sequences <- read.csv("Updater/March_2026/Output_data/UNITE_sequences_all_LAT.csv", row.names = "X")

# Remove locked sequences
UNITE_sequences <- UNITE_sequences %>%
  filter(!grepl("Locked", sequence, ignore.case = TRUE))

# Fix accession
# UNITE_sequences$Accession_Number <- str_split_i(UNITE_sequences$Accession_Number, "\\|", 1)

table(UNITE_sequences$Country)

# Just European Countries
euro_countries <- c("Russia", 'Germany', "United Kingdom", "France", "Italy", "Spain", "Poland", "Ukraine", "Romania", "Netherlands", "Belgium", "Czech Republic", "Sweden", "Portugal", "Greece", "Hungary", "Austria", "Belarus", "Switzerland", "Bulgaria", "Serbia", "Denmark", "Finland", "Norway", "Slovakia", "Ireland", "Croatia", "Bosnia and Herzegovina", "Moldova", "Lithuania", "Albania", "Slovenia", "Latvia", "North Macedonia", "Estonia", "Luxembourg", "Montenegro", "Malta", "Iceland", "Andorra", "Liechtenstein", "Monaco", "San Marino", "Holy See")

pattern <- str_c(euro_countries, collapse = "|")

UNITE_sequences_euro <- UNITE_sequences %>%
  mutate(Country = str_extract(Country, pattern)) %>%
  filter(!is.na(Country))

# Fix names
UNITE_sequences_euro <- UNITE_sequences_euro %>% dplyr::rename(species = Taxon_Name)

UNITE_sequences_euro <- fix_names(UNITE_sequences_euro, names)

# Parse Accession number
parse_unite_id <- function(x) {
  parts <- str_split(x, "\\|")[[1]] |> str_trim()
  
  is_udb <- str_detect(parts, "^UDB\\d+$")
  is_sh  <- str_detect(parts, "^SH\\d+(\\.\\d+FU)?$")
  is_is  <- str_detect(parts, "^IS\\d+$")
  is_insdc <- str_detect(parts, "^[A-Z]{1,4}\\d{5,}$") & !is_udb & !is_sh & !is_is
  
  tibble(
    original = x,
    unite_sample_id = if (any(is_udb)) parts[is_udb][1] else NA_character_,
    insdc_sample_id = if (any(is_insdc)) parts[is_insdc][1] else NA_character_,
    species_hypothesis = if (any(is_sh | is_is)) parts[is_sh | is_is][1] else NA_character_
  )
}

parsed_ids <- map_dfr(UNITE_sequences_euro$Accession_Number, parse_unite_id)

UNITE_sequences_euro <- bind_cols(UNITE_sequences_euro, parsed_ids %>% select(-original))

UNITE_sequences_euro <- UNITE_sequences_euro %>% select(-species_hypothesis)

# write.csv(UNITE_sequences_euro, "Updater/March_2026/Output_data/UNITE_sequences_euro_LAT.csv", row.names = F)

```

1474 current species with UNITE marker

50.5%

## BOLD
*First go to bold website and download all metadata for Ascomycota*
```{r}
library(httr2)
library(jsonlite)
library(dplyr)
library(purrr)
library(tibble)
library(BOLDconnectR)
library(httr2)
library(jsonlite)

bold.apikey("EA1F210E-F2A6-4512-AD76-38B0879ECC61")

# load in bold results
bold_asco_meta <- read.delim("Updater/March_2026/Input_data/bold_ascomycota_result_2026-03-13.tsv")

# Make clean species column and take process id and species
bold_asco_meta$clean_species <- sapply(strsplit(bold_asco_meta$species, "\\s+"), function(words) {
  paste(words[1:min(2, length(words))], collapse = " ")
})

# Load BLS names in list
species_list <- sort(unique(c(names$old_name, names$new_name)))

# Format query for BOLD
bold_query <- bold_asco_meta %>% select(processid, clean_species) %>% filter(clean_species %in% species_list)

# Use BOLDConnectR to query for processids
BCDM_data<-bold.fetch(get_by = "processid",
                      identifiers = bold_query$processid)

BCDM_data_summary<-bold.data.summarize(bold_df = BCDM_data,
                               summary_type = "concise_summary")

BCDM_data_summary$concise_summary

# Fix names
BCDM_data$species <- sapply(strsplit(BCDM_data$species, "\\s+"), function(words) {
  paste(words[1:min(2, length(words))], collapse = " ")
})

BCDM_data <- BCDM_data %>% dplyr::rename(sequence = nuc)

BCDM_data <- fix_names(BCDM_data, names)

# Reverse geocode to get country names for those with coords but no country.ocean
subset_data <- BCDM_data[is.na(BCDM_data$country.ocean) & !is.na(BCDM_data$coord), ]

coords_split <- do.call(rbind, strsplit(subset_data$coord, ","))

subset_data$lat <- as.numeric(coords_split[,1])
subset_data$lon <- as.numeric(coords_split[,2])

points_sf <- st_as_sf(subset_data, coords = c("lon", "lat"), crs = 4326)

countries <- ne_countries(scale = "medium", returnclass = "sf")

joined <- st_join(points_sf, countries["name"])

subset_data$country_filled <- joined$name

BCDM_data$country.ocean[
  is.na(BCDM_data$country.ocean) & !is.na(BCDM_data$coord)
] <- subset_data$country_filled

# Filter by country
BCDM_data <- BCDM_data %>% filter(country.ocean %in% euro_countries)

# Write out BOLD data
# write.csv(BCDM_data, "Updater/March_2026/Output_data/BOLD_sequences_all_LAT.csv")

# Write out taxids for later
taxids <- bold %>% select(taxid, species) %>% unique()

#write.csv(taxids, "Updater/March_2026/Output_data/ncbi_tax_ids.csv")

```

903 species

## Load and Merge sequence data
### BOLD and UNITE
```{r}
# BOLD
bold <- read.csv( "Updater/March_2026/Output_data/BOLD_sequences_all_LAT.csv", row.names = "X")
bold$Database <- "BOLD"

bold <- bold %>% select(sampleid, species, marker_code, sequence, country.ocean, Database) %>% dplyr::rename(country = country.ocean, database = Database, marker = marker_code)

# UNITE
unite <- read.csv("Updater/March_2026/Output_data/UNITE_sequences_euro_LAT.csv")

unite <- unite %>%
  mutate(sampleid = coalesce(insdc_sample_id, unite_sample_id))

unite <- unite %>% select(sampleid, species, Marker, sequence, Country, database) %>% rename(country = Country, marker = Marker)

# All data
all_sequences <- bind_rows(unite, bold)
all_sequences$sequence <- trimws(all_sequences$sequence)

# Make new marker names
all_sequences <- all_sequences %>%
  mutate(
    sequence = trimws(sequence),
    seq_length = nchar(sequence)
  ) %>%
  filter(!is.na(sequence), sequence != "", seq_length > 0) %>%
  mutate(
    marker_group = case_when(
      
      # keep mtSSU separate first so it doesn't get caught by SSU
      str_detect(marker, regex("^mtSSU$", ignore_case = TRUE)) ~ "mtSSU",
      
      # exact ITS1 / ITS2
      str_detect(marker, regex("^ITS1$", ignore_case = TRUE)) ~ "ITS1",
      str_detect(marker, regex("^ITS2$", ignore_case = TRUE)) ~ "ITS2",
      
      # explicit full ITS permutations (ITS1 + 5.8S + ITS2 only)
      str_detect(
        marker,
        regex("^(5\\.8S, ITS1, ITS2|5\\.8S, ITS2, ITS1|ITS1, 5\\.8S, ITS2|ITS1, ITS2, 5\\.8S|ITS2, 5\\.8S, ITS1|ITS2, ITS1, 5\\.8S)$",
              ignore_case = TRUE)
      ) ~ "ITS_full",
      
      # plain ITS with plausible full-length size
      str_detect(marker, regex("^ITS$", ignore_case = TRUE)) &
        seq_length >= 350 & seq_length <= 900 ~ "ITS_full",
      
      # ITS-containing markers that also include LSU and/or SSU = rOperon
      str_detect(marker, regex("ITS", ignore_case = TRUE)) &
        str_detect(marker, regex("LSU|28S|SSU|18S", ignore_case = TRUE)) ~ "rOperon",
      
      # plain ITS but unusually long = likely operon-scale
      str_detect(marker, regex("^ITS$", ignore_case = TRUE)) &
        seq_length > 900 ~ "rOperon",
      
      # standalone LSU
      str_detect(marker, regex("^(LSU rDNA|nucLSU|28S)$", ignore_case = TRUE)) ~ "LSU",
      
      # standalone SSU
      str_detect(marker, regex("^(SSU rDNA|18S)$", ignore_case = TRUE)) ~ "SSU",
      
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(marker_group))

# Then rename marker and remove group
all_sequences <- all_sequences %>% select(-marker, -seq_length) %>% dplyr::rename(marker = marker_group)

# Replace all ambiguous characters with 'N'
all_sequences$sequence <- gsub("[^ATCGNatcg]", "N", all_sequences$sequence)

# Remove parentheses and their contents from the headers
all_sequences$sampleid <- gsub("\\(.*\\)", "", all_sequences$sampleid)

# Trim any extra spaces that might have been introduced
all_sequences$sampleid <- trimws(all_sequences$sampleid)

```

### Add Martin7
```{r}
# Read the FASTA file
martin <- readDNAStringSet("Databases/Martin7/Martin7_ITS.fas")

# Read in metadata
martin_meta <- read_excel("Databases/Martin7/Martin7_metadata.xlsx")

# Extract headers and sequences
martin_df <- data.frame(headers = names(martin), sequences = as.character(martin))

# Reset rownames
row.names(martin_df) <- NULL

# split unique ID and species name
martin_df$sampleid <- str_split_i(martin_df$headers, "_", -1)

# Remove those that are just to genus
martin_df <- filter(martin_df, !grepl("sp.", headers))
martin_df <- filter(martin_df, !grepl("unidentified", headers))

# Remove aff., cf., unidentified,  and other odd characters for simplicity
martin_df$headers <- gsub("_cf.|_aff.|_(sensu_Sipman)|_TYPE1|_TYPE2|_TYPE3|_2|_var._flavicans|_var._isidiata|_var._muscorum", "", martin_df$headers)

# Create species column
martin_df$species <- paste0(str_split_i(martin_df$headers, "_", 1), " ", str_split_i(martin_df$headers, "_", 2))

# Add NR_ to beginning of these
martin_df[c(1086, 877, 1271, 829, 288, 294, 1170), 3] <- paste0("NR_", martin_df[c(1086, 877, 1271, 829, 288, 294, 1170), 3])

# Filter for UK species
martin_df <- unique(rbind(martin_df %>% filter(species %in% names$old_name), martin_df %>% filter(species %in% names$new_name)))

martin_df %>% dplyr::filter(species %in% names$old_name)

# Check if there are missing Sample IDs in sequence file (T = 173, F = 824)
table(martin_df$sampleid %in% all_sequences$sampleid)

# Join metadata for country
martin_df <- martin_df %>% left_join(select(martin_meta, `ITS accession`, Origin), by = join_by(sampleid == `ITS accession`))

# Remove those without country
martin_df <- na.omit(martin_df)

# Replace with consistent country names
martin_df$Origin <- gsub(".*Netherlands.*", "Netherlands", martin_df$Origin)
martin_df$Origin <- gsub(".*Russia.*", "Russia", martin_df$Origin)
martin_df$Origin <- gsub(".*Austria.*", "Austria", martin_df$Origin)
martin_df$Origin <- gsub(".*Sweden.*", "Sweden", martin_df$Origin)
martin_df$Origin <- gsub(".*Norway.*", "Norway", martin_df$Origin)
martin_df$Origin <- gsub(".*Great Britain.*", "United Kingdom", martin_df$Origin)
martin_df$Origin <- gsub(".*Ukraine.*", "Ukraine", martin_df$Origin)
martin_df$Origin <- gsub(".*Poland.*", "Poland", martin_df$Origin)
martin_df$Origin <- gsub(".*Greece.*", "Greece", martin_df$Origin)
martin_df$Origin <- gsub(".*USA.*|Hawaii", "USA", martin_df$Origin)
martin_df$Origin <- gsub(".*Korea.*", "Korea", martin_df$Origin)
martin_df$Origin <- gsub(".*Japan.*", "Japan", martin_df$Origin)
martin_df$Origin <- gsub(".*Spain.*", "Spain", martin_df$Origin)
martin_df$Origin <- gsub(".*Portugal.*", "Portugal", martin_df$Origin)
martin_df$Origin <- gsub(".*China.*", "China", martin_df$Origin)
martin_df$Origin <- gsub(".*Canada.*", "Canada", martin_df$Origin)
martin_df$Origin <- gsub("N/A|NA", NA, martin_df$Origin)
martin_df <- na.omit(martin_df)

# Select only european countries
martin_df <- martin_df %>%
  filter(Origin %in% euro_countries)

martin_df <- martin_df %>% dplyr::rename(sequence = sequences, country = Origin) %>% select(-headers)

martin_df <- fix_names(martin_df, names)

# Look at species
table(martin_df$species %in% all_sequences$species)
# 12 new species

# Join to other dataset
martin_df$marker <- "ITS_full"
martin_df$database <- "Martin7"

all_sequences <- rbind(all_sequences, martin_df)

```

1648 species

### Summarize and create rOperon DB DF
```{r}
# Filter just rOperon
table(all_sequences$marker)

its_seq <- filter(all_sequences, marker == "ITS_full" | marker == "ITS1" | marker == "ITS2")

# fix names
its_seq <- unique(fix_names(its_seq, names))

# take top 100 records
its_seq <- its_seq %>%
  mutate(Sequence_Length = nchar(sequence))

species_counts <- its_seq %>%
  count(species, name = "Occurrences")

# Identify species with more than 100 records
species_to_downsample <- species_counts %>%
  filter(Occurrences > 50) %>%
  pull(species)

# Keep 100 longest sequences for species with >50 records
its_seq <- its_seq %>%
  group_by(species) %>%
  filter(species %in% species_to_downsample) %>%              # Select overrepresented species
  slice_max(Sequence_Length, n = 50, with_ties = FALSE) %>%  # Keep top 100 longest sequences
  ungroup() %>%
  bind_rows(                                                  # Add back species with <=100 records
    its_seq %>% filter(!(species %in% species_to_downsample))
  ) %>% select(-Sequence_Length)

# Generate unique 6-digit codes
set.seed(123)  # Set seed for reproducibility
num_codes <- nrow(its_seq)
its_seq$unique_code <- paste0("BL", sprintf("%06d", sample(1e5:999999, num_codes, replace = FALSE)))

its_seq$Species_code <- gsub(" ", "_", its_seq$species)

its_seq$header <- paste(its_seq$unique_code, its_seq$sampleid, its_seq$Species_code, sep = "|")

# Create the DNAStringSet object
dna_sequences <- DNAStringSet(its_seq$sequence)

# Assign the names (headers) to the sequences
names(dna_sequences) <- its_seq$header

# Write the sequences to a FASTA file
# writeXStringSet(dna_sequences, filepath="Updater/March_2026/Output_data/BLITS_ITS.fasta", format="fasta")

```

#### Write and read in ITS
```{r}
# write.csv(its_seq, "Databases/BLITS/Final/its_seq_full.csv")

its_seq <- read.csv("Databases/BLITS/Final/its_seq_full.csv", row.names = "X")

```

Summary stats
1,470 species in rOperon DB

# nucLSU
## GenBank nucLSU
```{r}
library(gbfetch)
library(dplyr)

all_metadata_euro <- read.csv("Updater/March_2026/Output_data/genbank_meta_euro_marker.csv", row.names = "X")

nucLSU <- all_metadata_euro %>% filter(nucLSU == 1)

ncbi_api_key = "3250605898efe21223d0a656ad86a363c608"
Sys.setenv(ENTREZ_KEY = ncbi_api_key)

convert_to_bases <- function(numeric_sequence) {
  # Define the mapping vector
  mapping <- c("18" = "A", "28" = "C", "48" = "G", "88" = "T")
  # Convert numeric codes to characters using the mapping
  bases <- sapply(numeric_sequence, function(x) mapping[as.character(x)])
  # Collapse into a single string if desired
  sequence_string <- paste(bases, collapse = "")
  return(sequence_string)
}

nucLSU_seq <- data.frame()
convert_to_bases <- function(numeric_sequence) {
  mapping <- c("18" = "A", "28" = "C", "48" = "G", "88" = "T")
  bases <- unname(mapping[as.character(numeric_sequence)])
  paste(bases, collapse = "")
}

library(rentrez)
library(ape)
library(dplyr)

fetch_one_fasta <- function(acc) {
  if (is.na(acc) || acc == "") return(NA_character_)

  txt <- rentrez::entrez_fetch(
    db = "nucleotide",
    id = acc,
    rettype = "fasta",
    retmode = "text"
  )

  if (is.null(txt) || !nzchar(txt)) return(NA_character_)

  tf <- tempfile(fileext = ".fa")
  writeLines(txt, tf)
  on.exit(unlink(tf), add = TRUE)

  # Read as character, not DNAbin
  seqs <- ape::read.dna(
    file = tf,
    format = "fasta",
    as.character = TRUE
  )

  # seqs is usually a matrix with one row per sequence
  seq_string <- paste(seqs[1, ], collapse = "")
  seq_string
}

results <- vector("list", nrow(nucLSU))

for (i in seq_len(nrow(nucLSU))) {
  acc <- nucLSU$accession[i]

  results[[i]] <- tryCatch({
    seq_string <- fetch_one_fasta(acc)

    data.frame(
      accession = acc,
      species   = nucLSU$species[i],
      sequence  = seq_string,
      country   = nucLSU$country[i],
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    message("Failed at row ", i, " (", acc, "): ", conditionMessage(e))
    NULL
  })

  cat(i, "of", nrow(nucLSU), "\n")
}

nucLSU_seq <- bind_rows(results)
nucLSU_gb <- nucLSU_seq
#nucLSU_gb <- read.csv("Databases/genbank/nucLSU_seq_genbank.csv", row.names = "X")

nucLSU_gb$species <- gsub("\\]", "", nucLSU_gb$species)
nucLSU_gb$species <- gsub("\\[", "", nucLSU_gb$species)

nucLSU_gb$Database <- "GenBank"

# Name Replace
nucLSU_gb <- nucLSU_gb %>% filter(species %in% unique(names$old_name, names$new_name))
nucLSU_gb <- unique(fix_names(nucLSU_gb, names))

# Remove unrecoverable records
nucLSU_gb <- na.omit(nucLSU_gb)

# Generate unique 6-digit codes
set.seed(123)  # Set seed for reproducibility
num_codes <- nrow(nucLSU_gb)
nucLSU_gb$unique_code <- paste0("BL", sprintf("%06d", sample(1e5:999999, num_codes, replace = FALSE)))

nucLSU_gb$Species_code <- gsub(" ", "_", nucLSU_gb$species)

nucLSU_gb$header <- paste(nucLSU_gb$unique_code, nucLSU_gb$Species_code, sep = "|")

# Create the DNAStringSet object
dna_sequences <- DNAStringSet(nucLSU_gb$sequence)

# Assign the names (headers) to the sequences
names(dna_sequences) <- nucLSU_gb$header

#write.csv(nucLSU_gb, "Updater/March_2026/Output_data/nucLSU_seq_genbank.csv")
# Write the sequences to a FASTA file
writeXStringSet(dna_sequences, filepath="Updater/March_2026/Output_data/nucLSU_seq_genbank_nucLSU.fasta", format="fasta")

nuclsu_comb <- nucLSU_gb

```

# nucSSU
## GenBank mtSSU
```{r}

all_metadata_euro <- read.csv("Updater/March_2026/Output_data/genbank_meta_euro_marker.csv", row.names = "X")
nucSSU <- all_metadata_euro %>% filter(nucSSU == 1)
ncbi_api_key = "3250605898efe21223d0a656ad86a363c608"
Sys.setenv(ENTREZ_KEY = ncbi_api_key)

for (i in seq_len(nrow(nucSSU))) {
  acc <- nucSSU$accession[i]

  results[[i]] <- tryCatch({
    seq_string <- fetch_one_fasta(acc)

    data.frame(
      accession = acc,
      species   = nucSSU$species[i],
      sequence  = seq_string,
      country   = nucSSU$country[i],
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    message("Failed at row ", i, " (", acc, "): ", conditionMessage(e))
    NULL
  })

  cat(i, "of", nrow(nucSSU), "\n")
}

nucSSU_seq <- bind_rows(results)

# write.csv(nucSSU_seq, "Databases/genbank/nucSSU_seq_genbank.csv")
#nucSSU_seq <- read.csv("Databases/genbank/nucSSU_seq_genbank.csv", row.names = "X")

# Fix weird characters
nucSSU_seq$species <- gsub("\\]", "", nucSSU_seq$species)
nucSSU_seq$species <- gsub("\\[", "", nucSSU_seq$species)

nucSSU_seq$Database <- "GenBank"

# Name Replace
nucSSU_seq <- nucSSU_seq %>% filter(species %in% unique(names$old_name, names$new_name))
nucSSU_seq <- unique(fix_names(nucSSU_seq, names))
nucSSU_seq <- na.omit(nucSSU_seq)

# Generate unique 6-digit codes
set.seed(123)  # Set seed for reproducibility
num_codes <- nrow(nucSSU_seq)
nucSSU_seq$unique_code <- paste0("BL", sprintf("%06d", sample(1e5:999999, num_codes, replace = FALSE)))

nucSSU_seq$Species_code <- gsub(" ", "_", nucSSU_seq$species)

nucSSU_seq$header <- paste(nucSSU_seq$unique_code, nucSSU_seq$Species_code, sep = "|")

# Create the DNAStringSet object
dna_sequences <- DNAStringSet(nucSSU_seq$sequence)

# Assign the names (headers) to the sequences
names(dna_sequences) <- nucSSU_seq$header

# Write the sequences to a FASTA file
writeXStringSet(dna_sequences, filepath="Updater/March_2026/Output_data/nucSSU_seq_genbank.fasta", format="fasta")

```

# mtSSU
## GenBank mtSSU
```{r}

mtSSU <- all_metadata_euro %>% filter(mtSSU == 1)

results <- vector("list", nrow(mtSSU))

for (i in seq_len(nrow(mtSSU))) {
  acc <- mtSSU$accession[i]

  results[[i]] <- tryCatch({
    seq_string <- fetch_one_fasta(acc)

    data.frame(
      accession = acc,
      species   = mtSSU$species[i],
      sequence  = seq_string,
      country   = mtSSU$country[i],
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    message("Failed at row ", i, " (", acc, "): ", conditionMessage(e))
    NULL
  })

  cat(i, "of", nrow(mtSSU), "\n")
}

mtSSU_seq <- bind_rows(results)

# Fix weird characters
mtSSU_seq$species <- gsub("\\]", "", mtSSU_seq$species)
mtSSU_seq$species <- gsub("\\[", "", mtSSU_seq$species)

mtSSU_seq$Database <- "GenBank"

# Name Replace
mtSSU_seq <- mtSSU_seq %>% filter(species %in% unique(names$old_name, names$new_name))
mtSSU_seq <- unique(fix_names(mtSSU_seq, names))
mtSSU_seq <- na.omit(mtSSU_seq)

# Generate unique 6-digit codes
set.seed(123)  # Set seed for reproducibility
num_codes <- nrow(mtSSU_seq)
mtSSU_seq$unique_code <- paste0("BL", sprintf("%06d", sample(1e5:999999, num_codes, replace = FALSE)))

mtSSU_seq$Species_code <- gsub(" ", "_", mtSSU_seq$species)

mtSSU_seq$header <- paste(mtSSU_seq$unique_code, mtSSU_seq$Species_code, sep = "|")

# Create the DNAStringSet object
dna_sequences <- DNAStringSet(mtSSU_seq$sequence)

# Assign the names (headers) to the sequences
names(dna_sequences) <- mtSSU_seq$header

# Write the sequences to a FASTA file
writeXStringSet(dna_sequences, filepath="Updater/March_2026/Output_data/mtSSU_seq_genbank.fasta", format="fasta")

write.csv(mtSSU_seq, "Updater/March_2026/Output_data/mtSSU_seq_genbank.csv")

```

## Martin7
```{r}
# Read the FASTA file
martin <- readDNAStringSet("Databases/Martin7/Martin7_mtSSU.fas")

# Read in metadata
martin_meta <- read_excel("Databases/Martin7/Martin7_metadata.xlsx")

# Extract headers and sequences
martin_df <- data.frame(headers = names(martin), sequences = as.character(martin))

# Reset rownames
row.names(martin_df) <- NULL

# split unique ID and species name
martin_df$SampleID <- str_split_i(martin_df$headers, "_", -1)

# Remove those that are just to genus
martin_df <- filter(martin_df, !grepl("sp.", headers))
martin_df <- filter(martin_df, !grepl("unidentified", headers))

# Remove aff., cf., unidentified,  and other odd characters for simplicity
martin_df$headers <- gsub("_cf.|_aff.|_(sensu_Sipman)|_TYPE1|_TYPE2|_TYPE3|_2|_var._flavicans|_var._isidiata|_var._muscorum", "", martin_df$headers)

martin_df <- martin_df %>%
  filter(str_count(headers, "_") != 1)

# Create species column
martin_df$Species <- paste0(str_split_i(martin_df$headers, "_", 1), " ", str_split_i(martin_df$headers, "_", 2))

# Filter for UK species
martin_df <- unique(rbind(martin_df %>% filter(Species %in% names$old_name), martin_df %>% filter(Species %in% names$new_name)))

# Join metadata for country
martin_df <- martin_df %>% left_join(select(martin_meta, `mtSSU accession`, Origin), by = join_by(SampleID == `mtSSU accession`))

# Remove those without country
martin_df <- na.omit(martin_df)

# Replace with consistent country names
martin_df$Origin <- gsub(".*Netherlands.*", "Netherlands", martin_df$Origin)
martin_df$Origin <- gsub(".*Russia.*", "Russia", martin_df$Origin)
martin_df$Origin <- gsub(".*Austria.*", "Austria", martin_df$Origin)
martin_df$Origin <- gsub(".*Sweden.*", "Sweden", martin_df$Origin)
martin_df$Origin <- gsub(".*Norway.*", "Norway", martin_df$Origin)
martin_df$Origin <- gsub(".*Great Britain.*", "United Kingdom", martin_df$Origin)
martin_df$Origin <- gsub(".*Ukraine.*", "Ukraine", martin_df$Origin)
martin_df$Origin <- gsub(".*Poland.*", "Poland", martin_df$Origin)
martin_df$Origin <- gsub(".*Greece.*", "Greece", martin_df$Origin)
martin_df$Origin <- gsub(".*USA.*|Hawaii", "USA", martin_df$Origin)
martin_df$Origin <- gsub(".*Korea.*", "Korea", martin_df$Origin)
martin_df$Origin <- gsub(".*Japan.*", "Japan", martin_df$Origin)
martin_df$Origin <- gsub(".*Spain.*", "Spain", martin_df$Origin)
martin_df$Origin <- gsub(".*Portugal.*", "Portugal", martin_df$Origin)
martin_df$Origin <- gsub(".*China.*", "China", martin_df$Origin)
martin_df$Origin <- gsub(".*Canada.*", "Canada", martin_df$Origin)
martin_df$Origin <- gsub("N/A|NA", NA, martin_df$Origin)
martin_df <- na.omit(martin_df)

# Just European Countries
euro_countries <- c("Russia", 'Germany', "United Kingdom", "France", "Italy", "Spain", "Poland", "Ukraine", "Romania", "Netherlands", "Belgium", "Czech Republic", "Sweden", "Portugal", "Greece", "Hungary", "Austria", "Belarus", "Switzerland", "Bulgaria", "Serbia", "Denmark", "Finland", "Norway", "Slovakia", "Ireland", "Croatia", "Bosnia and Herzegovina", "Moldova", "Lithuania", "Albania", "Slovenia", "Latvia", "North Macedonia", "Estonia", "Luxembourg", "Montenegro", "Malta", "Iceland", "Andorra", "Liechtenstein", "Monaco", "San Marino", "Holy See")

# Select only european countries
martin_df <- martin_df %>%
  filter(Origin %in% euro_countries)

# Look at species
#table(martin_df$Species %in% all_sequences$Species)
# 207 new species

# Join to other dataset
martin_df$Marker <- "mtSSU"
martin_df$Database <- "Martin7"
martin_df <- martin_df %>% dplyr::rename(sequence = sequences, species = Species, country = Origin) %>% select(-headers, -SampleID, -Marker, -Database)

martin_df$Database <- "Martin7"

```

### Combine mtSSU
```{r}
mtSSU_gb <- read.csv("Databases/genbank/mtSSU_seq_genbank.csv", row.names = "X")

# Fix weird characters
mtSSU_gb$species <- gsub("\\]", "", mtSSU_gb$species)
mtSSU_gb$species <- gsub("\\[", "", mtSSU_gb$species)

mtSSU_gb$Database <- "GenBank"

# Bind data
mtssu_comb <- mtSSU_gb %>% select(-accession) %>% 
  bind_rows(martin_df)

# Name Replace
mtssu_comb <- mtssu_comb %>% filter(species %in% unique(names$old_name, names$new_name))
mtssu_comb <- unique(fix_names(mtssu_comb, names))
mtssu_comb <- na.omit(mtssu_comb)

# Generate unique 6-digit codes
set.seed(123)  # Set seed for reproducibility
num_codes <- nrow(mtssu_comb)
mtssu_comb$unique_code <- paste0("BL", sprintf("%06d", sample(1e5:999999, num_codes, replace = FALSE)))

mtssu_comb$Species_code <- gsub(" ", "_", mtssu_comb$species)

mtssu_comb$header <- paste(mtssu_comb$unique_code, mtssu_comb$Species_code, sep = "|")

# Create the DNAStringSet object
dna_sequences <- DNAStringSet(mtssu_comb$sequence)

# Assign the names (headers) to the sequences
names(dna_sequences) <- mtssu_comb$header

# Write the sequences to a FASTA file
# writeXStringSet(dna_sequences, filepath="Databases/BLITS/Final/BLRD_mtssu.fasta", format="fasta")

```


# Create Tax File
```{r}

ncbi_api_key = ""
Sys.setenv(ENTREZ_KEY = ncbi_api_key)

# Example: Retrieve taxonomy for a list of species names
species_names <- unique(c(names$new_name))
databases <- c("ncbi", "itis", "gbif", "col")

# Initialize an empty list to store results
taxonomy_results <- list()

species_names <- data.frame(species_names)

#species_names <- species_names %>% filter(species_names %in% rerun$full_ITS_species_list)

# Loop over each species
for (species in species_names$species_names) {
  found <- FALSE
  
  # Loop through each database until a match is found
  for (db in databases) {
    # Try to get the classification
    result <- tryCatch({
      if (db == "ncbi") {
        classification(species, db = db, key = ncbi_api_key, rows = 1)
      } else {
        classification(species, db = db, rows = 1)
      }
    }, error = function(e) NULL)
    
    # Check if result is valid and not NULL or an atomic vector
    if (is.list(result) && !is.null(result[[1]]) && "name" %in% names(result[[1]])) {
      # Store the result and mark it as found
      taxonomy_results[[species]] <- data.frame(Species = species, 
                                                Taxonomy = paste(result[[1]]$name, collapse = "; "), 
                                                Database = db)
      found <- TRUE
      break  # Exit the inner loop once a match is found
    }
  }
  
  # If not found in any database, add a 'Not found' entry
  if (!found) {
    taxonomy_results[[species]] <- data.frame(Species = species, Taxonomy = NA, Database = "Not found")
  }
}

# Combine results into a single data frame
taxonomy_results_df <- do.call(rbind, taxonomy_results)

# write.csv(taxonomy_results_df, "taxonomy_all_british_lichen.csv")

```

## Manual curation of tax file
```{r}

# Read taxonomy files
myco_tax <- read.csv("Taxonomy/taxonomy_all_british_lichen.csv")
photo_tax <- read.csv("Taxonomy/taxonomy_chlorophyta.csv")

# Myco
# Pre-allocate the dataframe with the correct number of rows
tax_df <- data.frame(
  kingdom = character(nrow(myco_tax)),
  phylum = character(nrow(myco_tax)),
  class = character(nrow(myco_tax)),
  order = character(nrow(myco_tax)),
  family = character(nrow(myco_tax)),
  genus = character(nrow(myco_tax)),
  species = character(nrow(myco_tax)),
  stringsAsFactors = FALSE
)

# Loop through each row of myco_tax
for (i in 1:nrow(myco_tax)) {
  
  # Print the progress
  print(i)
  
  # Extract components
  components <- strsplit(myco_tax$Taxonomy[i], ";")[[1]] %>% trimws()
  
  # Extract groups based on the rules, assigning NA if no match is found
  tax_df$kingdom[i] <- "Fungi" # Always "Fungi"
  tax_df$phylum[i] <- ifelse(length(components[grepl("mycota$", components)]) > 0, 
                             components[grepl("mycota$", components)][1], NA)
  tax_df$class[i] <- ifelse(length(components[grepl("mycetes$", components)]) > 0, 
                            components[grepl("mycetes$", components)][1], NA)
  tax_df$order[i] <- ifelse(length(components[grepl("ales$", components)]) > 0, 
                            components[grepl("ales$", components)][1], NA)
  tax_df$family[i] <- ifelse(length(components[grepl("aceae$", components)]) > 0, 
                             components[grepl("aceae$", components)][1], NA)
  tax_df$genus[i] <- ifelse(length(components) >= 2, components[length(components) - 1], NA)
  tax_df$species[i] <- ifelse(length(components) >= 1, components[length(components)], NA)
}

# Manual curation of several that didn't load
tax_df[321, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Pertusariales", "Megasporaceae", "Circinaria", "Circinaria tuberculosa")
tax_df[1267, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Ostropales", "Graphidaceae", "Graphis", "Graphis persoonii")
tax_df[1327, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Rhizocarpales", "Rhizocarpaceae", "Haugania", "Haugania oederi")
tax_df[1408, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Ostropales", "Stictidaceae", "Karstenia", "Karstenia chrysophaea")
tax_df[1409, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Ostropales", "Stictidaceae", "Karstenia", "Karstenia dictyospora")
tax_df[1410, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Ostropales", "Stictidaceae", "Karstenia", "Karstenia nigra")
tax_df[1410, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Ostropales", "Stictidaceae", "Karstenia", "Karstenia nigra")
tax_df[2320, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", NA, NA, "Randlanea", "Randlanea usneicola")
tax_df[1188, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Lecanorales", "Pilocarpaceae", "Epicladonia", "Epicladonia simplex")
tax_df[1189, ] <- c("Fungi", "Ascomycota", "Lecanoromycetes", "Lecanorales", "Pilocarpaceae", "Epicladonia", "Epicladonia stenospora")
tax_df[2516, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Asterotexales", NA, "Taeniolella", "Taeniolella punctata")
tax_df[2518, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Asterotexales", NA, "Taeniolella", "Taeniolella toruloides")
tax_df[58, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Monoblastiales", "Monoblastiaceae", "Acrocordia", "Acrocordia cavata")
tax_df[59, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Monoblastiales", "Monoblastiaceae", "Acrocordia", "Acrocordia conoidea")
tax_df[63, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Monoblastiales", "Monoblastiaceae", "Acrocordia", "Acrocordia subglobosa")
tax_df[64, ] <- c("Fungi", "Ascomycota", "Dothideomycetes", "Asterinales", "Stictographaceae", "Actinocladium", "Actinocladium rhodosporum")
tax_df[206, 5] <- "Chrysothricaceae"
tax_df[241, 3:5] <- c("Dothideomycetes", "Pleosporales", "Arthopyreniaceae")

# Algae and cyanobacteria
tax_df <- tax_df[-c(1707, 1712, 1718), ]

# Create the mapping table from the original tax_df
genus_mapping <- tax_df %>%
  select(kingdom, phylum, class, order, family, genus) %>%
  distinct()

genus_mapping <- na.omit(genus_mapping)
genus_mapping$family <- gsub("Byssolomataceae", "Pilocarpaceae", genus_mapping$family)

# Manually add in some missing
Fulgensia <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Lecanoromycetes", order = "Teloschistales", family = "Teloschistaceae", genus = "Fulgensia")
Secoliga <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Lecanoromycetes", order = "Gyalectales", family = "Gyalectaceae", genus = "Secoliga")
Stictographa <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Dothideomycetes", order = "Asterotexales", family = "Stictographaceae", genus = "Stictographa")
Burgoa <- data.frame(kingdom = "Fungi", phylum = "Basidiomycota", class = "Agaricomycetes", order = "Cantharellales", family = "Hydnaceae", genus = "Burgoa")
Illosporiopsis <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Sordariomycetes", order = "Hypocreales", family = "Hypocreales_family_incertae_sedis", genus = "Illosporiopsis")
Bachmanniomyces <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Sordariomycetes", order = "Pezizomycotina_order_incertae_sedis", family = "Pezizomycotina_family_incertae_sedis", genus = "Bachmanniomyces")
Hemigrapha <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Dothideomycetes", order = "Asterinales", family = "Hemigraphaceae", genus = "Hemigrapha")
Corticifraga <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Lecanoromycetes", order = "Lecanorales", family = "Lecanoromycetes_family_incertae_sedis", genus = "Corticifraga")
Clathroporinopsis <- data.frame(kingdom = "Fungi", phylum = "Ascomycota", class = "Lecanoromycetes", order = "Gyalectales", family = "Gyalectaceae", genus = "Clathroporinopsis")

genus_mapping <- bind_rows(genus_mapping, Fulgensia, Secoliga, Stictographa, Burgoa, Illosporiopsis, Bachmanniomyces, Hemigrapha, Corticifraga, Clathroporinopsis)

# Identify genera mapped to multiple families
conflicting_genus <- genus_mapping %>%
  group_by(genus) %>%
  filter(n_distinct(family) > 1) %>%
  arrange(genus)

# Reset row names to match sequential indices
genus_mapping <- genus_mapping %>% mutate(row_id = row_number())
rownames(genus_mapping) <- NULL

# Fix manually
# Combine all row indices into a vector
rows_to_remove <- c(92, 216, 224, 140, 387, 494, 134, 191, 524, 553, 
                    523, 438, 505, 370, 28, 448, 182, 595, 219)

# Remove all specified rows at once
genus_mapping <- genus_mapping[-rows_to_remove, ]

# Fix species names
tax_df$sequence <- "test"
tax_df <- fix_names(tax_df, names)
tax_df <- tax_df %>% select(-sequence)

# Manually add in some missing
missing_sp <- data.frame(kingdom = NA, phylum = NA, class =  NA, order = NA, family = NA, genus = NA, species = unique(names$new_name))

# Bind into tax_df
tax_df <- bind_rows(tax_df, missing_sp)

# Extract genus from the corrected species column
tax_df <- tax_df %>%
  mutate(genus = str_split_i(species, " ", 1))

# Standardize genus column in both datasets
tax_df <- tax_df %>%
  mutate(genus = str_trim(genus))

# Create the genus_mapping from tax_df
genus_mapping_2 <- tax_df %>%
  select(kingdom, phylum, class, order, family, genus) %>%
  group_by(genus) %>%
  summarise(
    kingdom = head(na.omit(kingdom), 1),
    phylum = head(na.omit(phylum), 1),
    class = head(na.omit(class), 1),
    order = head(na.omit(order), 1),
    family = head(na.omit(family), 1),
    .groups = "drop"
  )

# Step 2: Combine and ensure mapping is complete
genus_mapping <- unique(bind_rows(genus_mapping, genus_mapping_2))

genus_mapping <- genus_mapping %>%
  mutate(genus = str_trim(genus))

# Step 4: Join the mapping table to update all ranks
tax_df <- tax_df %>%
  left_join(genus_mapping, by = "genus", suffix = c("", "_updated"), relationship = "many-to-many") %>%
  mutate(
    kingdom = coalesce(kingdom_updated, kingdom),
    phylum = coalesce(phylum_updated, phylum),
    class = coalesce(class_updated, class),
    order = coalesce(order_updated, order),
    family = coalesce(family_updated, family)
  ) %>%
  select(-ends_with("_updated"))

tax_df$family <- gsub("Byssolomataceae", "Pilocarpaceae", tax_df$family)

tax_df <- tax_df %>% select(-row_id)
tax_df <- unique(tax_df)
tax_df <- tax_df %>% filter(!is.na(family))

# Find genera in tax_df that are not in genus_mapping
unmatched_genera <- tax_df %>%
  filter(is.na(kingdom) | is.na(phylum) | is.na(class) | is.na(order) | is.na(family)) %>%
  distinct(genus)

print(unmatched_genera)

entries_to_remove <- tibble(
  kingdom = c(
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi","Fungi",
    "Fungi","Fungi","Fungi","Fungi","Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi", "Fungi"
  ),
  phylum = c(
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota",
    "Ascomycota","Ascomycota","Ascomycota","Ascomycota","Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota", "Ascomycota"
  ),
  class = c(
    "Dothideomycetes","Dothideomycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes","Dothideomycetes",
    "Eurotiomycetes","Eurotiomycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes",
    "Lecanoromycetes","Lecanoromycetes","Lecanoromycetes","Dothideomycetes","Lecanoromycetes","Lecanoromycetes",
    "Lecanoromycetes","Lecanoromycetes","Eurotiomycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes",
    "Lecanoromycetes","Lecanoromycetes","Lecanoromycetes","Dothideomycetes","Lecanoromycetes","Lecanoromycetes", "Lecanoromycetes","Lichinomycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes", "Lecanoromycetes","Lecanoromycetes","Dothideomycetes","Lecanoromycetes","Lecanoromycetes","Lecanoromycetes", "Lecanoromycetes","Dothideomycetes", "Dothideomycetes", "Eurotiomycetes", "Dothideomycetes", "Lecanoromycetes", "Lecanoromycetes", "Lichinomycetes", "Lecanoromycetes", "Dothideomycetes", "Leotiomycetes", "Leotiomycetes", "Lecanoromycetes", "Eurotiomycetes", "Lecanoromycetes", "Dothideomycetes", "Eurotiomycetes", "Dothideomycetes", "Dothideomycetes"
  ),
  order = c(
    "Eremithallales","Pleosporales","Gyalectales","Teloschistales","Lecideales","Pleosporales",
    "Sclerococcales","Sclerococcales","Gyalectales","Ostropales","Ostropales","Teloschistales",
    "Teloschistales","Lecanorales","Ostropales","Lichenoconiales","Lecanorales","Lecanorales",
    "Lecanorales","Lecanorales","Verrucariales","Lecanorales","Pertusariales","Gyalectales",
    "Teloschistales","Teloschistales","Trapeliales","Pleosporales","Pertusariales","Lecanorales",
    "Lichinales","Gyalectales","Teloschistales","Schaereriales","Ostropales","Sclerococcales",
    "Lecanorales","Lecanorales","Capnodiales","Lecanorales","Caliciales","Lecanorales","Ostropales","Collemopsidiales", "Mycosphaerellales","Sclerococcales", "Collemopsidiales", "Lecanorales", "Teloschistales", "Lichinales", "Gyalectales", "Pleosporales", "Lecanorales", "Leotiales", "Lecideales", "Chaetothyriales", "Lecanorales", "Asterinales", "Verrucariales", "
Dothideomycetes_order_incertae_sedis", "Asterotexales"
  ),
  family = c(
    "Melaspileaceae","Arthopyreniaceae","Gyalectaceae","Physciaceae","Lecideaceae","Arthopyreniaceae",
    "Dactylosporaceae","Dactylosporaceae","Gyalectaceae","Graphidaceae","Graphidaceae","Leprocaulaceae",
    "Physciaceae","Hymeneliaceae","Graphidaceae","Lichenoconiaceae","Ectolechiaceae","Sarrameanaceae",
    "Micareaceae","Helocarpaceae","Verrucariaceae","Ramalinaceae","Pertusariaceae","Gyalectaceae",
    "Physciaceae","Physciaceae","Trapeliaceae","Dacampiaceae","Pertusariaceae","Sphaerophoraceae",
    "Lichinaceae","Gyalectaceae","Physciaceae","Schaereriaceae","Sagiolechiaceae","Sclerococcaceae",
    "Lecanoraceae","Pilocarpaceae","Mycosphaerellaceae","Lecanoraceae","Physciaceae","Vezdaeaceae",
    "Graphidaceae","Xanthopyreniaceae", "Mycosphaerellaceae", "Sclerococcaceae", "Xanthopyreniaceae", "Catillariaceae", "Physciaceae", "Lichinaceae", "Gyalectaceae", "Abrothallaceae", "Pilocarpaceae", "Mniaeciaceae", "Lecideaceae", "Verrucariaceae", "Melaspileaceae", "Melaspileaceae", "Verrucariaceae", "Pyrenidiaceae", "Mytilinidiaceae"
  ),
  genus = c(
    "Arthonia","Arthopyrenia","Belonia","Buellia","Calvitimela","Collemopsidium",
    "Dactylospora","Dactylospora","Gyalecta","Gyalidea","Gyalideopsis","Halecania",
    "Heterodermia","Ionaspis","Jamesiella","Lichenoconium","Lopadium","Loxospora",
    "Micarea","Micarea","Muellerella","Mycobilimbia","Ochrolechia","Pachyphiale",
    "Physcia","Physconia","Placynthiella","Polycoccum","Porina","Protomicarea",
    "Protothelenella","Ramonia","Rinodina","Ropalospora","Sagedia","Sclerococcum",
    "Scoliciosporum","Scutula","Stigmidium","Strangospora","Tetramelas","Vezdaea",
    "Xerotrema","Zwackhiomyces", "Arthopyrenia", "Sclerococcum", "Collemopsidium", "Halecania", "Diplotomma", "Protothelenella", "Ramonia", "Abrothallus", "Epicladonia", "Epithamnolia", "Lecania", "Lichenodiplis", "Melaspilea", "Melaspilea", "Phaeospora", "Pyrenidium", "Taeniolella"
  )
)

# Remove rows matching any of these entries
tax_df <- tax_df %>%
  anti_join(entries_to_remove, by = c("kingdom", "phylum", "class", "order", "family", "genus"))

genus_mapping <- genus_mapping %>%
  anti_join(entries_to_remove, by = c("kingdom", "phylum", "class", "order", "family", "genus"))

tax_df <- unique(tax_df)

tax_df <- tax_df %>% filter(species %in% unique(names$new_name))

unique(names %>% filter(!new_name %in% unique(tax_df$species)) %>% select(new_name))

tax_df <- tax_df %>%
  mutate(order = case_when(
    family == "Protothelenellaceae" ~ "Baeomycetales",
    TRUE ~ order   # leave all other values of `order` unchanged
  ))

# Extract rows with duplicated species names
duplicated_species_df <- tax_df %>%
  group_by(species) %>%
  filter(n() > 1) %>%
  ungroup()

# View the resulting dataframe
duplicated_species_df

unique(names %>% filter(!str_split_i(names$new_name, " ", 1) %in% unique(tax_df$genus)) %>% select(new_name))

genera <- unique(data.frame(genus = str_split_i(names$new_name, " ", 1)))
genera <- genera %>% filter(!genus %in% tax_df$genus)

# stats
table(tax_df$phylum) / nrow(tax_df)
# 2.2% of British licehn are basidiomycetes and 97.8% are ascos

length(unique(tax_df$class))
sort(table(tax_df$class)/ nrow(tax_df))
# 16 classes

# Write out final file
# write.csv(tax_df, "Taxonomy/taxonomy_british_lichen_final.csv")

```

## Manual additions from Consortium of Lichen Herbaria Taxonomy Explorer
```{r}
# Manually compiled taxonomy for 20 genera (best current estimates).

genus_taxonomy <- tibble::tribble(
  ~genus,             ~kingdom, ~phylum,      ~class,             ~order,           ~family,

  # 1) Acremonium
  "Acremonium",       "Fungi",  "Ascomycota", "Sordariomycetes",  "Hypocreales",    "Hypocreaceae",

  # 2) Roselliniella
  "Roselliniella",    "Fungi",  "Ascomycota", "Sordariomycetes",  "Sordariales",  "Sordariales_family_incertae_sedis",

  # 3) Peridothelia
  "Peridothelia",     "Fungi",  "Ascomycota", "Eurotiomycetes",   "Verrucariales",  "Verrucariaceae",

  # 4) Adelolechia
  "Adelolechia",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",    "Lecideaceae",

  # 5) Corticiruptor
  "Corticiruptor",    "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanoromycetes_order_incertae_sedis",               "Lecanoromycetes_family_incertae_sedis",

  # 6) Arthophacopsis
  "Arthophacopsis",   "Fungi",  "Ascomycota", "Dothideomycetes",  "Abrothallales",  "Abrothallaceae",

  # 7) Mycoglaena
  "Mycoglaena",       "Fungi",  "Ascomycota", "Dothideomycetes",  "Pleosporales",   "Arthopyreniaceae",

  # 8) Crustodiplodina
  "Crustodiplodina",  "Fungi",  "Ascomycota", "Dothideomycetes",  NA,               NA,

  # 9) Schaereria
  "Schaereria",       "Fungi",  "Ascomycota", "Lecanoromycetes",  "Schaereriales",  "Schaereriaceae",

  # 10) Bachmanniomyces
  "Bachmanniomyces",  "Fungi",  "Ascomycota", "Pezizomycotina_class_incertae_sedis",   "Pezizomycotina_order_incertae_sedis",               "Pezizomycotina_family_incertae_sedis",

  # 11) Aquacida
  "Aquacida",         "Fungi",  "Ascomycota", NA, NA, NA,

  # 12) Bactrospora
  "Bactrospora",      "Fungi",  "Ascomycota", "Arthoniomycetes",  "Arthoniales",    "Roccellaceae",

  # 13) Clathroporinopsis
  "Clathroporinopsis","Fungi",  "Ascomycota", "Lecanoromycetes",  "Gyalectales",    "Gyalectaceae",

  # 14) Myochroidea
  "Myochroidea",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Peltigerales",   "Pannariaceae",

  # 15) Puttea
  "Puttea",           "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",    "Tephromelataceae",

  # 16) Sarcosagium
  "Sarcosagium",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Baeomycetales",  "Baeomycetaceae",

  # 17) Biatoridium
  "Biatoridium",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",    "Pezizomycotina_family_incertae_sedis",

  # 18) Piccolia
  "Piccolia",         "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",    "Piccoliaceae",

  # 19) Lecanidion
  "Lecanidion",       "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",    "Pilocarpaceae",

  # 20) Intralichen
  "Intralichen",      "Fungi",  "Ascomycota", "Dothideomycetes",  "Pezizomycotina_order_incertae_sedis",               "Pezizomycotina_family_incertae_sedis",

  # 21) Botryolepraria
  "Botryolepraria", "Fungi",  "Ascomycota",    "Lecanoromycetes",    "Leprocaulales",      "Leprocaulaceae",

  # 22) Buelliella
  "Buelliella",     "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 23) Burgoa
  "Burgoa",         "Fungi",  "Basidiomycota", "Agaricomycetes",     "Cantharellales",     "Hydnaceae",

  # 24) Leightoniomyces
  "Leightoniomyces","Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 25) Cercidospora
  "Cercidospora",   "Fungi",  "Ascomycota",    "Dothideomycetes",    "Pleosporales",       "Dacampiaceae",

  # 26) Chionosphaera
  "Chionosphaera",  "Fungi",  "Basidiomycota", "Cystobasidiomycetes","Cystobasidiales",    "Cystobasidiaceae",

  # 27) Cladoniicola
  "Cladoniicola",   "Fungi",  "Ascomycota",    "Leotiomycetes",      NA,                   NA,

  # 28) Lathagium
  "Lathagium",      "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 29) Milospium
  "Milospium",      "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 30) Corticifraga
  "Corticifraga",   "Fungi",  "Ascomycota",    "Eurotiomycetes",     "Sclerococcales",     "Dactylosporaceae",

  # 31) Cyrtidula
  "Cyrtidula",      "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 32) Lichenosticta
  "Lichenosticta",  "Fungi",  "Ascomycota",    "Dothideomycetes",    "Abrothallales",      "Abrothallaceae",

  # 33) Endococcus
  "Endococcus",     "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   "Endococcaceae",

  # 34) Neoechinodiscus
  "Neoechinodiscus","Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 35) Dermatocrpon (often spelled Dermatocarpon)
  "Dermatocrpon",   "Fungi",  "Ascomycota",    "Eurotiomycetes",     "Verrucariales",      "Verrucariaceae",

  # 36) Epiphloea
  "Epiphloea",      "Fungi",  "Ascomycota",    "Eurotiomycetes",     NA,                   NA,

  # 37) Everniicola
  "Everniicola",    "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 38) Feltgeniomyces
  "Feltgeniomyces", "Fungi",  "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 39) Fulgensis (often placed near Fulgensia)
  "Fulgensis",      "Fungi",  "Ascomycota",    "Lecanoromycetes",    "Teloschistales",     "Teloschistaceae",

  # 40) Hemigrapha
  "Hemigrapha",     "Fungi",  "Ascomycota",    "Dothideomycetes",    "Asterinales",        "Hemigraphaceae",
  
  # 41) Illosporiopsis
  # Commonly placed in Sordariomycetes with uncertain order/family (lichenicolous)
  "Illosporiopsis", "Fungi",          "Ascomycota",    "Sordariomycetes",    NA,                    NA,

  # 42) Homostegia
  # Often lichenicolous, tentatively in Pleosporales (Dacampiaceae) or incertae sedis
  "Homostegia",     "Fungi",          "Ascomycota",    "Dothideomycetes",    "Pleosporales",       "Dacampiaceae",

  # 43) Hymenobia
  # A lichen genus in Lichinomycetes (sometimes in Lichinaceae)
  "Hymenobia",      "Fungi",          "Ascomycota",    "Lichinomycetes",     "Lichinales",         "Lichinaceae",

  # 44) Illosporium
  # Related to Illosporiopsis, generally Sordariomycetes incertae sedis
  "Illosporium",    "Fungi",          "Ascomycota",    "Sordariomycetes",    NA,                    NA,

  # 45) Kalchbrenneriella
  # Lichenicolous fungus often placed in Dothideomycetes incertae sedis
  "Kalchbrenneriella","Fungi",        "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 46) Lecidoma
  # e.g. Lecidoma demissum, placed in Lecideales, Lecideaceae
  "Lecidoma",       "Fungi",          "Ascomycota",    "Lecanoromycetes",    "Lecideales",         "Lecideaceae",

  # 47) Sporostatia
  # Typically in Lecanoromycetes, Rhizocarpales; some refs place it in Sporastatiaceae or Megalariaceae
  "Sporostatia",    "Fungi",          "Ascomycota",    "Lecanoromycetes",    "Rhizocarpales",      "Sporastatiaceae",

  # 48) Scytonema
  # A genus of cyanobacteria (photobiont), not a fungus
  "Scytonema",      "Bacteria",       "Cyanobacteria", "Cyanophyceae",       "Nostocales",         "Scytonemataceae",

  # 49) Desmococcus
  # A genus of green algae (photobiont) in the Chlorophyta
  "Desmococcus",    "Plantae",        "Chlorophyta",   "Trebouxiophyceae",   NA,                    NA,

  # 50) Trentepohlia
  # A filamentous green alga in Ulvophyceae (Trentepohliales, Trentepohliaceae)
  "Trentepohlia",   "Plantae",        "Chlorophyta",   "Ulvophyceae",        "Trentepohliales",    "Trentepohliaceae",

  # 51) Monodictys
  # Dematiaceous hyphomycete, commonly Dothideomycetes incertae sedis
  "Monodictys",     "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 52) Lichenohendersonia
  # Lichenicolous fungus, typically Dothideomycetes incertae sedis
  "Lichenohendersonia","Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 53) Lichenopuccinia
  # A basidiomycetous, lichenicolous fungus, likely in Pucciniomycetes
  "Lichenopuccinia","Fungi",          "Basidiomycota", "Pucciniomycetes",    "Pucciniales",        NA,

  # 54) Melaspileopsis
  # Some references place it in Dothideomycetes, Asterinales
  "Melaspileopsis", "Fungi",          "Ascomycota",    "Dothideomycetes",    "Asterinales",        NA,

  # 55) Stictographa
  # Typically in Dothideomycetes, Asterotexales, Stictographaceae
  "Stictographa",   "Fungi",          "Ascomycota",    "Dothideomycetes",    "Asterotexales",      "Stictographaceae",

  # 56) Minutoexcipula
  # Lichenicolous fungus in Dothideomycetes; order/family often uncertain
  "Minutoexcipula", "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 57) Minutophoma
  # Another lichenicolous genus, typically Dothideomycetes incertae sedis
  "Minutophoma",    "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 58) Mixtoconidium
  # Lichenicolous fungus in Ascomycota, often Dothideomycetes incertae sedis
  "Mixtoconidium",  "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 59) Neolamya
  # e.g. Neolamya peltigerae, also Dothideomycetes incertae sedis
  "Neolamya",       "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,

  # 60) Nigromacula
  # e.g. Nigromacula uniseptata, typically Ascomycota incertae sedis or Dothideomycetes incertae sedis
  "Nigromacula",    "Fungi",          "Ascomycota",    "Dothideomycetes",    NA,                    NA,
  
  # 61) arthonia
  # Genus of lichenized/lichenicolous fungi in class Arthoniomycetes
  "arthonia",       "Fungi",       "Ascomycota",    "Arthoniomycetes",    "Arthoniales",        "Arthoniaceae",

  # 62) Roselliniopsis
  # Generally placed in Dothideomycetes incertae sedis or near Pleosporales
  "Roselliniopsis", "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 63) Cucurbidothis
  # Likely Dothideomycetes incertae sedis (rarely documented lichenicolous genus)
  "Cucurbidothis",  "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 64) Vouauxiomyces
  # Lichenicolous, typically in Dothideomycetes incertae sedis
  "Vouauxiomyces",  "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 65) Bastenia
  # Another rarely placed lichenicolous genus, often left in Dothideomycetes incertae sedis
  "Bastenia",       "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 66) Degelia
  # A lichen genus in Lecanoromycetes, order Peltigerales, family Pannariaceae
  "Degelia",        "Fungi",       "Ascomycota",    "Lecanoromycetes",    "Peltigerales",       "Pannariaceae",

  # 67) Sarcogyna
  # Possibly allied with Sarcogyne (Acarosporaceae) or another uncertain group. 
  # We'll place in Lecanoromycetes, Acarosporales, Acarosporaceae as a best guess.
  "Sarcogyna",      "Fungi",       "Ascomycota",    "Lecanoromycetes",    "Acarosporales",      "Acarosporaceae",

  # 68) Strigula
  # Well-known in Dothideomycetes, Strigulales, Strigulaceae
  "Strigula",       "Fungi",       "Ascomycota",    "Dothideomycetes",    "Strigulales",        "Strigulaceae",

  # 69) Psammina
  # Lichenicolous genus in Dothideomycetes, often placed in Pleosporales, Dacampiaceae
  "Psammina",       "Fungi",       "Ascomycota",    "Dothideomycetes",    "Pleosporales",       "Dacampiaceae",

  # 70) Pseudorobillarda
  # Dematiaceous fungus, typically Dothideomycetes incertae sedis
  "Pseudorobillarda","Fungi",      "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 71) Psorotichia
  # Lichen genus in Lichinomycetes, Lichinales, Lichinaceae
  "Psorotichia",    "Fungi",       "Ascomycota",    "Lichinomycetes",     "Lichinales",         "Lichinaceae",

  # 72) Raciborskiomyces
  # Lichenicolous fungus in Dothideomycetes incertae sedis
  "Raciborskiomyces","Fungi",      "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 73) Randlanea
  # Known from Randlanea usneicola (lichenicolous), placed in Lecanoromycetes incertae sedis
  "Randlanea",      "Fungi",       "Ascomycota",    "Lecanoromycetes",    NA,                   NA,

  # 74) Reconditella
  # Lichenicolous, typically Dothideomycetes incertae sedis
  "Reconditella",   "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 75) Refractohilum
  # Another lichenicolous genus, Dothideomycetes incertae sedis
  "Refractohilum",  "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 76) Rhaphidicyrtis
  # Lichenicolous fungus in Dothideomycetes incertae sedis
  "Rhaphidicyrtis", "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 77) Roesleria
  # Known from Roesleria subterranea, in class Leotiomycetes, order Helotiales, family Roesleriaceae
  "Roesleria",      "Fungi",       "Ascomycota",    "Leotiomycetes",      "Helotiales",         "Roesleriaceae",

  # 78) Rosellinula
  # Lichenicolous in Dothideomycetes incertae sedis
  "Rosellinula",    "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 79) Sarcopyrenia
  # Eurotiomycetes, order Sarcopyreniales, family Sarcopyreniaceae
  "Sarcopyrenia",   "Fungi",       "Ascomycota",    "Eurotiomycetes",     "Sarcopyreniales",    "Sarcopyreniaceae",

  # 80) Vouauxiella
  # Lichenicolous, typically Dothideomycetes incertae sedis
  "Vouauxiella",    "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 81) Sporormiella
  # Placed in Dothideomycetes, Pleosporales, family Sporormiaceae
  "Sporormiella",   "Fungi",       "Ascomycota",    "Dothideomycetes",    "Pleosporales",       "Sporormiaceae",

  # 82) Talpapellis
  # Lichenicolous fungus in Dothideomycetes incertae sedis
  "Talpapellis",    "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 83) Taeniolina
  # Dematiaceous fungus in Dothideomycetes incertae sedis
  "Taeniolina",     "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 84) Trichoconis
  # Dematiaceous hyphomycete, generally Dothideomycetes incertae sedis
  "Trichoconis",    "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 85) Trichothecium
  # Class Sordariomycetes, order Hypocreales, family Trichotheciaceae
  "Trichothecium",  "Fungi",       "Ascomycota",    "Sordariomycetes",    "Hypocreales",        "Trichotheciaceae",

  # 86) Naetrcymbe
  # Lichenicolous fungus in Dothideomycetes incertae sedis
  "Naetrcymbe",     "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA,

  # 87) Zevadia
  # Lichenicolous fungus in Dothideomycetes incertae sedis
  "Zevadia",        "Fungi",       "Ascomycota",    "Dothideomycetes",    NA,                   NA
)

# 1) Fill missing `order` with "_order_incertae_sedis", climbing up the ranks
genus_taxonomy <- genus_taxonomy %>%
  mutate(
    order = case_when(
      !is.na(order) ~ order,
      !is.na(class) ~ paste0(class, "_order_incertae_sedis"),
      !is.na(phylum) ~ paste0(phylum, "_order_incertae_sedis"),
      !is.na(kingdom) ~ paste0(kingdom, "_order_incertae_sedis"),
      TRUE ~ NA_character_
    )
  )

# 2) Fill missing `family`. 
#    If `order` ends with "_order_incertae_sedis", replace only that suffix with "_family_incertae_sedis".
#    Otherwise, do the usual logic of appending "_family_incertae_sedis".
missing_genera <- genus_taxonomy %>%
  mutate(
    family = case_when(
      !is.na(family) ~ family,
      
      # If `order` is something like "Ascomycota_order_incertae_sedis"
      # then do a string replace so we don't double-incertae_sedis
      !is.na(order) & str_detect(order, "_order_incertae_sedis$") ~ str_replace(
        order, "_order_incertae_sedis$", "_family_incertae_sedis"
      ),
      
      # Else if `order` is non-NA but doesn't have that suffix
      !is.na(order) ~ paste0(order, "_family_incertae_sedis"),
      
      # If `order` is NA, climb up to `class`, `phylum`, then `kingdom`
      !is.na(class) ~ paste0(class, "_family_incertae_sedis"),
      !is.na(phylum) ~ paste0(phylum, "_family_incertae_sedis"),
      !is.na(kingdom) ~ paste0(kingdom, "_family_incertae_sedis"),
      TRUE ~ NA_character_
    )
  )

missing_genera[11,4] <- "Ascomycota_class_incertae_sedis"

names_w_genera <- data.frame(genus = str_split_i(names$new_name, " ", 1), species = names$new_name)

missing_genera <- missing_genera %>% left_join(names_w_genera)

tax_df_2 <- bind_rows(tax_df, missing_genera)
missing <- unique(tax_df_2 %>% filter(!species %in% unique(names$new_name)))

tax_df <-tax_df_2

tax_df <- tax_df %>% filter(kingdom == "Fungi")

tax_df_missing <- tax_df %>% filter(!species == unique(names$new_name))

missing <- names %>% filter(!new_name %in% tax_df$species)
unique(missing$new_name)

missing_tax <- tibble::tribble(
  ~kingdom, ~phylum, ~class, ~order, ~family, ~genus, ~species,

  # 1) Acremonium
  "Fungi",  "Ascomycota", "Lecanoromycetes",  NA, "Epigloeaceae", "Epigloea", "Epigloea soleiformis")

tax_df <- tax_df %>% filter(!species == unique(missing$new_name))

tax_df <- bind_rows(tax_df, missing_tax)

length(unique(tax_df$species))

# write.csv(tax_df, "C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Taxonomy/taxonomy_british_lichen_final.csv")


```

### Calculate Taxonomy presence
```{r}
tax_df_na <- tax_df
tax_df_na$class <- gsub("incertae_sedis", NA, tax_df$class)
tax_df_na$order <- gsub("incertae_sedis", NA, tax_df$order)
tax_df_na$family <- gsub("incertae_sedis", NA, tax_df$family)
tax_df_na$genus <- gsub("incertae_sedis", NA, tax_df$genus)

length(unique(tax_df_na$class))
length(unique(tax_df_na$order))
length(unique(tax_df_na$family))
length(unique(tax_df_na$genus))
length(unique(tax_df_na$species))

tax_df_na_its <- tax_df_na %>% filter(species %in% its_seq$species)

length(unique(tax_df_na_its$class))
length(unique(tax_df_na_its$order))
length(unique(tax_df_na_its$family))
length(unique(tax_df_na_its$genus))
length(unique(tax_df_na_its$species))


tax_df_na_nlsu <- tax_df_na %>% filter(species %in% nuclsu_comb$species)

length(unique(tax_df_na_nlsu$class))
length(unique(tax_df_na_nlsu$order))
length(unique(tax_df_na_nlsu$family))
length(unique(tax_df_na_nlsu$genus))
length(unique(tax_df_na_nlsu$species))

tax_df_na_nssu <- tax_df_na %>% filter(species %in% nucSSU_seq$species)

length(unique(tax_df_na_nssu$class))
length(unique(tax_df_na_nssu$order))
length(unique(tax_df_na_nssu$family))
length(unique(tax_df_na_nssu$genus))
length(unique(tax_df_na_nssu$species))

tax_df_na_mtssu <- tax_df_na %>% filter(species %in% mtssu_comb$species)

length(unique(tax_df_na_mtssu$class))
length(unique(tax_df_na_mtssu$order))
length(unique(tax_df_na_mtssu$family))
length(unique(tax_df_na_mtssu$genus))
length(unique(tax_df_na_mtssu$species))

tax_df_na_all <- unique(rbind(tax_df_na_its, tax_df_na_nlsu, tax_df_na_nssu, tax_df_na_mtssu))

length(unique(tax_df_na_all$class))
length(unique(tax_df_na_all$order))
length(unique(tax_df_na_all$family))
length(unique(tax_df_na_all$genus))
length(unique(tax_df_na_all$species))

```

## Taxonomy Updater
```{r}

tax_df <- read.csv("C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Taxonomy/taxonomy_british_lichen_final.csv")

new_tax_query <- names %>% filter(!new_name %in% tax_df$species) %>% select(new_name) %>% unique()

ncbi_api_key = "3250605898efe21223d0a656ad86a363c608"
Sys.setenv(ENTREZ_KEY = ncbi_api_key)

# Example: Retrieve taxonomy for a list of species names
databases <- c("ncbi", "itis", "gbif", "col")

# Initialize an empty list to store results
taxonomy_results <- list()

# Loop over each species
for (species in new_tax_query$new_name) {
  found <- FALSE
  
  # Loop through each database until a match is found
  for (db in databases) {
    # Try to get the classification
    result <- tryCatch({
      if (db == "ncbi") {
        classification(species, db = db, key = ncbi_api_key, rows = 1)
      } else {
        classification(species, db = db, rows = 1)
      }
    }, error = function(e) NULL)
    
    # Check if result is valid and not NULL or an atomic vector
    if (is.list(result) && !is.null(result[[1]]) && "name" %in% names(result[[1]])) {
      # Store the result and mark it as found
      taxonomy_results[[species]] <- data.frame(Species = species, 
                                                Taxonomy = paste(result[[1]]$name, collapse = "; "), 
                                                Database = db)
      found <- TRUE
      break  # Exit the inner loop once a match is found
    }
  }
  
  # If not found in any database, add a 'Not found' entry
  if (!found) {
    taxonomy_results[[species]] <- data.frame(Species = species, Taxonomy = NA, Database = "Not found")
  }
}

# Combine results into a single data frame
taxonomy_results_df <- do.call(rbind, taxonomy_results)

# Myco
# Pre-allocate the dataframe with the correct number of rows
tax_df_new <- data.frame(
  kingdom = character(nrow(taxonomy_results_df)),
  phylum = character(nrow(taxonomy_results_df)),
  class = character(nrow(taxonomy_results_df)),
  order = character(nrow(taxonomy_results_df)),
  family = character(nrow(taxonomy_results_df)),
  genus = character(nrow(taxonomy_results_df)),
  species = character(nrow(taxonomy_results_df)),
  stringsAsFactors = FALSE
)

# Loop through each row of myco_tax
for (i in 1:nrow(taxonomy_results_df)) {
  
  # Print the progress
  print(i)
  
  # Extract components
  components <- strsplit(taxonomy_results_df$Taxonomy[i], ";")[[1]] %>% trimws()
  
  # Extract groups based on the rules, assigning NA if no match is found
  tax_df_new$kingdom[i] <- "Fungi" # Always "Fungi"
  tax_df_new$phylum[i] <- ifelse(length(components[grepl("mycota$", components)]) > 0, 
                             components[grepl("mycota$", components)][1], NA)
  tax_df_new$class[i] <- ifelse(length(components[grepl("mycetes$", components)]) > 0, 
                            components[grepl("mycetes$", components)][1], NA)
  tax_df_new$order[i] <- ifelse(length(components[grepl("ales$", components)]) > 0, 
                            components[grepl("ales$", components)][1], NA)
  tax_df_new$family[i] <- ifelse(length(components[grepl("aceae$", components)]) > 0, 
                             components[grepl("aceae$", components)][1], NA)
  tax_df_new$genus[i] <- str_split_i(taxonomy_results_df$Species[i], " ", 1)
  tax_df_new$species[i] <- taxonomy_results_df$Species[i]
}

# Check which genera have already been placed
table(unique(tax_df_new$genus) %in% unique(tax_df$genus))

# columns you want to fill, but NOT genus/species
fill_cols <- c("kingdom", "phylum", "class", "order", "family")

# make a genus-level reference table from tax_df
genus_lookup <- tax_df %>%
  select(genus, all_of(fill_cols)) %>%
  distinct(genus, .keep_all = TRUE)

# join and fill only missing values in tax_df_new
tax_df_new_filled <- tax_df_new %>%
  left_join(genus_lookup, by = "genus", suffix = c("", "_ref")) %>%
  mutate(
    across(
      all_of(fill_cols),
      ~ coalesce(.x, get(paste0(cur_column(), "_ref")))
    )
  ) %>%
  select(-ends_with("_ref"))

# Now manually fill the rest
tax_df_new_filled_missing <- tax_df_new_filled  %>% filter(is.na(family) | is.na(order) | is.na(class))

# A few typos
tax_df_new_filled_missing[8,7] <- "Xanthomendoza fulva"
tax_df_new_filled_missing[8,6] <- "Xanthomendoza"

# manual curation
genus_taxonomy <- tibble::tribble(
  ~genus,               ~kingdom, ~phylum,      ~class,             ~order,            ~family,

  "Xanthocarpia",       "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Scythioria",         "Fungi",  "Ascomycota", "Dothideomycetes",  "Capnodiales",     "Mycosphaerellaceae",
  "Flavoplaca",         "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Kuettlingeria",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Chaenotricha",       "Fungi",  "Ascomycota", "Coniocybomycetes", "Coniocybales",    "Coniocybaceae",
  "Xanthomendoza",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Gyalectiphila",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Gyalectales",     "Gomphillaceae",
  "Lecanoropsis",       "Fungi",  "Ascomycota", "Lecanoromycetes",  "Lecanorales",     "Lecanoraceae",
  "Calogaya",           "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Lichenostella",      "Fungi",  "Ascomycota", "Lecanoromycetes",  "Arthoniales",     "Opegraphaceae",
  "Polycauliona",       "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Neobaryopsis",       "Fungi",  "Ascomycota", "Leotiomycetes",    "Helotiales",      "Dermateaceae",
  "Paraepicoccum",      "Fungi",  "Ascomycota", "Dothideomycetes",  "Pleosporales",    "Didymellaceae",
  "Huneckii",           "Fungi",  "Ascomycota", "Lecanoromycetes",  "Teloschistales",  "Teloschistaceae",
  "Pseudophaeoisaria",  "Fungi",  "Ascomycota", "Sordariomycetes",  "Xylariales",      "Xylariaceae",
  "Tingiopsidium",      "Fungi",  "Ascomycota", "Arthoniomycetes",  "Arthoniales",     "Arthoniaceae",
  "Venturia",           "Fungi",  "Ascomycota", "Dothideomycetes",  "Venturiales",     "Venturiaceae",
  "Xylohyphopsis",      "Fungi",  "Ascomycota", "Sordariomycetes",  "Xylariales",      "Xylariaceae",
  "Thelocarpon",  "Fungi",  "Ascomycota", NA, "Thelocarpales", "Thelocarpaceae"
)

# make new tax_df_completed
tax_df_new_filled_missing <- tax_df_new_filled_missing %>% select(genus, species) %>%
  left_join(genus_taxonomy, by = "genus")

# remove new names that don't exist anymore from tax_df
tax_df_remaining <- tax_df %>% filter(species %in% names$new_name) 

# Combine dataframes
tax_df_new_filled <- na.omit(tax_df_new_filled)

tax_df_new_2 <- rbind(tax_df_new_filled_missing, tax_df_new_filled)

tax_df_new_2 <- unique(tax_df_new_2)

tax_df_updated <- tax_df_new_2 %>% filter(is.na(family))

tax_df_updated <- rbind(tax_df_remaining, tax_df_new_2)

names %>% filter(!new_name %in% tax_df_updated$species)

# Write out
# write.csv(tax_df_updated, "Updater/March_2026/Output_data/taxonomy_british_lichen_final.csv", row.names = F)

```

# Read in data if previously retrieved
```{r}
# Write
#write.csv(mtssu_comb, "Databases/BLITS/Final/mtssu_comb.csv")
#write.csv(nuclsu_comb, "Databases/BLITS/Final/nuclsu_comb.csv")
#write.csv(nucSSU_seq, "Databases/BLITS/Final/nucssu_comb.csv")
#write.csv(its_seq, "Databases/BLITS/Final/its_seq_full.csv")

its_seq <- read.csv("Databases/BLITS/Final/its_seq_full.csv", row.names = "X")
mtssu_comb <- read.csv("Databases/BLITS/Final/mtssu_comb.csv", row.names = "X")
nuclsu_comb <- read.csv("Databases/BLITS/Final/nuclsu_comb.csv", row.names = "X")
nucSSU_seq <- read.csv("Databases/BLITS/Final/nucssu_comb.csv", row.names = "X")

tax_df <- read.csv("C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Taxonomy/taxonomy_british_lichen_final.csv")

# Combined Sequence Data
colnames(its_seq)
colnames(mtssu_comb)
colnames(nuclsu_comb)
colnames(nucSSU_seq)

mtssu_comb$Marker <- "mtSSU"
nuclsu_comb$Marker <- "nucLSU"
nucSSU_seq$Marker <- "nucSSU"

sequence_data_all_LAT <- bind_rows(its_seq, mtssu_comb, nuclsu_comb, nucSSU_seq)

write.csv(sequence_data_all_LAT, "C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Updater/March_2026/Output_data/sequence_data_all_LAT.csv", row.names = F)

```

# Figures

# Figure 1. Name changes over time
## Prep
```{r}
# Extract the 4-digit year from the text in parentheses, e.g. (1909), (1990), etc.
bls <- bls %>%
  mutate(
    Year = str_extract(Authority, "\\(\\d{4}\\)")        # Extract "(YYYY)"
  ) %>%
  mutate(
    Year = str_extract(Year, "\\d{4}") %>% as.integer()  # Convert to integer
  )

year_counts <- bls %>%
  filter(!is.na(Year)) %>%         # remove any rows without a valid Year
  group_by(Year) %>%
  summarize(count = n(), .groups = "drop")


year_counts[1,1] <- 2015
year_counts[2,1] <- 2018


# Updated Flora dataset
flora_df <- data.frame(
  Year = c(1801, 1871, 1921, 1965, 1970, 1992, 2009, 2015, 2017, 2018, 2022, 2023, 2025),
  Number_on_UK_checklist = c(225, 787, 1255, 1365, 1372, 1487, 1894, 1901, 1959, 1965, 2016, 2040, 2068),
  Flora = c("", "Leighton", "Annie Lorrain Smith", "Laundon", "Duncan", "Purvis et al.", "Smith et al.", "BLS", "BLS", "BLS", "BLS", "BLS", "BLS")
)

# Add the described species dataset
total_over_time <- bls %>% select(Name, Current.Taxon.Name, Is.Current.Name., Year)
total_over_time[6769, 4] <- 2015
total_over_time[7582, 4] <- 2018

df <- total_over_time %>%
  mutate(
    Current.Taxon.Name = if_else(
      is.na(Current.Taxon.Name) | Current.Taxon.Name == "",
      Name,
      Current.Taxon.Name
    ),
    FinalName = if_else(Is.Current.Name. == "Yes", Name, Current.Taxon.Name)
  )

df_simplified <- df %>%
  group_by(FinalName) %>%
  summarize(
    earliest_year = min(Year, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.infinite(earliest_year))

df_counts <- df_simplified %>%
  group_by(earliest_year) %>%
  summarize(n_new_species = n(), .groups = "drop")

all_years <- seq(min(df_counts$earliest_year), max(df_counts$earliest_year))

df_full <- data.frame(Year = all_years) %>%
  left_join(df_counts, by = c("Year" = "earliest_year")) %>%
  mutate(n_new_species = replace_na(n_new_species, 0)) %>%
  mutate(total_species = cumsum(n_new_species))

# Add the only lichen described species
total_over_time_l <- bls %>% dplyr::mutate(Group = replace_na(Group, "NA")) %>% dplyr::filter(!Group == "LF") %>%
  filter(!Group == "F") %>%
  filter(!Group == "CR") %>%
  select(Name, Current.Taxon.Name, Is.Current.Name., Year)

df_2 <- total_over_time_l %>%
  mutate(
    Current.Taxon.Name = if_else(
      is.na(Current.Taxon.Name) | Current.Taxon.Name == "",
      Name,
      Current.Taxon.Name
    ),
    FinalName = if_else(Is.Current.Name. == "Yes", Name, Current.Taxon.Name)
  )

df_2_simplified <- df_2 %>%
  group_by(FinalName) %>%
  summarize(
    earliest_year = min(Year, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.infinite(earliest_year))

df_2_counts <- df_2_simplified %>%
  group_by(earliest_year) %>%
  summarize(n_new_species = n(), .groups = "drop")

all_years_2 <- seq(min(df_2_counts$earliest_year), max(df_2_counts$earliest_year))

df_full_2 <- data.frame(Year = all_years_2) %>%
  left_join(df_2_counts, by = c("Year" = "earliest_year")) %>%
  mutate(n_new_species = replace_na(n_new_species, 0)) %>%
  mutate(total_species = cumsum(n_new_species))

# master text size
base_size <- 16
bls_pts <- flora_df %>% filter(Flora == "BLS")

# put the label a bit to the right of the last year
label_x <- max(bls_pts$Year) + 3
label_y <- bls_pts %>% filter(Year == max(Year)) %>% pull(Number_on_UK_checklist)

bls_label_df <- data.frame(x = label_x, y = label_y, lab = "BLS")

seg_df <- bls_pts %>%
  transmute(x    = Year,
            y    = Number_on_UK_checklist,
            xend = label_x,
            yend = label_y)

x_pad  <- 5                     
xmin   <- min(df_full$Year)
xmax   <- max(flora_df$Year) + x_pad

```

## Graph
*Fix B plot 0 marker above the x axis*
```{r}
library(grid)   # for unit()

base_size <- 16

lichen_theme <- theme_classic(base_size = base_size) +
  theme(
    # outer frame – thinner
    panel.border      = element_rect(color = "black", fill = NA, linewidth = 0.7),

    # keep only y-axis line; let panel border be the x-axis
    axis.line.x       = element_blank(),
    axis.line.y       = element_blank(),

    # ticks: thinner, black to match edges
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(6, "pt"),

    # labels: bold titles, internal labels (tick labels) plain + slightly larger, all black
    axis.title        = element_text(size = base_size + 2, face = "bold", colour = "black"),
    axis.text         = element_text(size = base_size + 1, face = "plain", colour = "black"),

    # legend – all black text
    legend.position   = "top",
    legend.title      = element_blank(),
    legend.text       = element_text(size = base_size - 2, colour = "black"),
    legend.key.width  = unit(18, "pt"),
    legend.key.height = unit(12, "pt")
  )

# --- 1) Shared x-range across BOTH panels ---
xmin <- min(c(df_full$Year, df_full_2$Year, flora_df$Year, year_counts$Year), na.rm = TRUE)
xmax <- max(c(df_full$Year, df_full_2$Year, flora_df$Year, year_counts$Year), na.rm = TRUE) + 2

# make the legend/y-label fonts a touch smaller than base
legend_sz <- base_size - 4
ytitle_sz <- base_size - 1

# --- 2) Panel B (bottom): cumulative curves, legend lives here ---
g1 <- ggplot() +
  geom_line(data = df_full,
            aes(Year, total_species, colour = "Described LAT Species"),
            linewidth = 0.7) +    # thinner trend
  geom_line(data = df_full_2,
            aes(Year, total_species, colour = "Described L Species"),
            linewidth = 0.7) +
  geom_line(data = flora_df,
            aes(Year, Number_on_UK_checklist, colour = "Number of Species in Flora"),
            linewidth = 0.7) +
  geom_point(data = flora_df,
             aes(Year, Number_on_UK_checklist,
                 colour = "Number of Species in Flora"),
             size = 4, stroke = 0.8) +
  scale_colour_manual(values = c(
    "Number of Species in Flora" = "grey10",
    "Described LAT Species"      = "grey60",
    "Described L Species"        = "grey80"
  )) +
  scale_x_continuous(limits = c(xmin, xmax), expand = c(0, 0)) +
  labs(x = "Year", y = "Number of species") +
  lichen_theme +
  theme(
    axis.title.y   = element_text(size = ytitle_sz, face = "bold", colour = "black"),
    axis.title.x   = element_text(size = ytitle_sz, face = "bold", colour = "black"),
    legend.position = "bottom",
    legend.text     = element_text(size = legend_sz, colour = "black"),
    legend.title    = element_blank(),
    plot.margin     = margin(6, 6, 6, 8)
  )

# --- 3) Panel A (top): bars + thinner loess curve, same x-range ---
ylim_max <- max(year_counts$count, na.rm = TRUE) * 1.05

g2 <- ggplot(year_counts, aes(Year, count)) +
  geom_col(fill = "grey85", colour = "grey55", linewidth = 0.4, width = 0.9) +
  geom_smooth(method = "loess", span = 0.4,
              colour = "black", linewidth = 0.6, se = FALSE) +  # thinner trend
  scale_y_continuous(limits = c(0, ylim_max), expand = c(0, 0)) +
  scale_x_continuous(limits = c(xmin, xmax), expand = c(0, 0)) +
  labs(x = NULL, y = "Described species per year") +
  lichen_theme +
  theme(
    axis.title.y   = element_text(size = ytitle_sz, face = "bold", colour = "black"),
    axis.text.x    = element_blank(),
    axis.ticks.x   = element_blank(),
    legend.position = "none",
    plot.margin     = margin(5, 5, 2, 8)
  )

# --- 4) Stack vertically with aligned x-axes & bottom legend ---
combined <- ggpubr::ggarrange(
  g2, g1,
  labels = c("A", "B"),
  font.label = list(size = base_size + 2, face = "bold", colour = "black"),
  ncol = 1, heights = c(1, 1), align = "v",
  common.legend = TRUE, legend = "bottom"
)

height_mm <- 200  # adjust if you want shorter/taller on the page

ggsave(
  filename = "Figures/Final Figures/Figure1_170mmV2.pdf",
  plot     = combined,
  device   = "pdf",
  width    = 170,
  height   = height_mm,
  units    = "mm",
  bg       = "white"
)


```


## Figure 2. Number of species per database
```{r}
plot_markers_bars_base <- function(dat) {
  stopifnot(all(c("Database","marker","count") %in% names(dat)))

  ## --------- tunables ---------
  y_max      <- max(dat$count, na.rm = TRUE) + 25
  bar_w      <- 0.95
  gap_w      <- 0.02   # smaller gap between bars within marker
  grp_gap    <- 0.70
  pad_l      <- 0.35
  pad_r      <- 0.25

  labs_cex   <- 0.75      # axis title size
  axis_cex   <- 0.55      # tick label size
  leg_cex    <- axis_cex  # legend text same as tick labels
  tick_col   <- "black"
  border_lwd <- 0.8       # frame + legend line width

  # ----- greys: darkest first (BOLD) -----
  db_lvls <- levels(dat$Database)
  greys   <- c(
    "BOLD"         = "grey0",
    "GenBank"      = "grey30",
    "Global Fungi" = "grey55",
    "Martin7"      = "grey75",
    "UNITE"        = "grey92"
  )
  greys   <- greys[db_lvls]
  markers <- levels(dat$marker)

  # widths of each marker group
  k_vec      <- sapply(markers, function(mk) nrow(subset(dat, marker == mk)))
  grp_widths <- k_vec * bar_w + pmax(k_vec - 1, 0) * gap_w

  x_left  <- c(pad_l, pad_l + cumsum(head(grp_widths + grp_gap, -1)))
  x_right <- x_left + grp_widths
  x_ctr   <- (x_left + x_right) / 2
  total_x <- tail(x_right, 1) + pad_r
  
  ## --------- set up empty plot ---------
  op <- par(
    mar = c(2.4, 2.4, 0.4, 0.4),    # a touch more left margin for y-title
    xaxs = "i",
    yaxs = "i",
    mgp  = c(1.2, 0.35, 0)         # same spacing logic for x & y
  )
  on.exit(par(op), add = TRUE)
  
  plot(NA, NA,
       xlim = c(0, total_x),
       ylim = c(0, y_max),
       xlab = "",
       ylab = "",
       xaxt = "n", yaxt = "n",
       bty  = "n")

  ## ---- axes (no axis line; we draw frame manually) ----
  axis(2, las = 1, cex.axis = axis_cex,
       col = tick_col, col.axis = tick_col, col.ticks = tick_col,
       tcl = -0.25, lwd = 0, lwd.ticks = 0.7)
  mtext("Number of species", side = 2, line = 1.4,
        cex = labs_cex, font = 2)

  axis(1, at = x_ctr, labels = markers, cex.axis = axis_cex,
       col = tick_col, col.axis = tick_col, col.ticks = tick_col,
       tcl = -0.25, lwd = 0, lwd.ticks = 0.7)
  mtext("Marker", side = 1, line = 1.4,
        cex = labs_cex, font = 2)

  ## ---- uniform frame (no thicker bit) ----
  segments(0,       0,       total_x, 0,       lwd = border_lwd)  # bottom
  segments(0,       0,       0,       y_max,   lwd = border_lwd)  # left
  segments(0,       y_max,   total_x, y_max,   lwd = border_lwd)  # top
  segments(total_x, 0,       total_x, y_max,   lwd = border_lwd)  # right

  ## --------- draw bars ---------
  for (i in seq_along(markers)) {
    mk <- markers[i]
    d  <- subset(dat, marker == mk)
    d  <- d[order(d$Database), , drop = FALSE]

    k <- nrow(d)
    if (k == 0) next

    width_i <- grp_widths[i]
    start_i <- x_left[i] +
      (width_i - (k * bar_w + (k - 1) * gap_w)) / 2
    x_lefts  <- start_i + (0:(k - 1)) * (bar_w + gap_w)
    x_rights <- x_lefts + bar_w

    for (j in seq_len(k)) {
      rect(x_lefts[j], 0, x_rights[j], d$count[j],
           col    = greys[as.character(d$Database[j])],
           border = "black", lwd = 0.7)
    }
  }

  ## --------- LEGEND INSET + L-SHAPE (BOTTOM + RIGHT) ---------
  # 1) compute layout
  lg <- legend("topleft",
               inset      = 0.015,
               x.intersp  = 0.5,   # a bit more horizontal spacing
               y.intersp  = 1.1,   # more vertical spacing between items
               bty        = "n",
               fill       = greys,
               border     = NA,
               legend     = db_lvls,
               ncol       = 1,
               cex        = leg_cex,
               title      = NULL,
               text.col   = "black",
               plot       = FALSE)

  # 2) draw legend
  legend(x      = lg$rect$left,
         y      = lg$rect$top + 25,
         legend = db_lvls,
         fill   = greys,
         border = NA,
         ncol   = 1,
         cex    = leg_cex,
         title  = NULL,
         text.col  = "black",
         bty    = "n",
         x.intersp = 0.5,
         y.intersp = 1.2)

  # 3) L-shaped box lines: extend just enough to meet axes
  lg_left   <- lg$rect$left
  lg_top    <- lg$rect$top
  lg_right  <- lg_left + lg$rect$w
  lg_bottom <- lg_top  - lg$rect$h

  ## bottom segment: from y-axis (x=0) to right edge of legend, at legend bottom
  segments(0, lg_bottom,
           lg_right, lg_bottom,
           lwd = border_lwd)

  ## right segment: from legend bottom up to top frame (y_max)
  segments(lg_right, lg_bottom,
           lg_right, y_max,
           lwd = border_lwd)
}

# EXPORT
pdf("Figures/Final Figures/Figure_2.pdf",
    width = 84/25.4,
    height = 90/25.4)

plot_markers_bars_base(markers_db)

dev.off()

```

## Figure 3. Number of species Venn Diagram
```{r}
venn.plot <- venn.diagram(
  x = list(
    ITS    = species_ITS,
    mtSSU  = species_mtSSU,
    nucLSU = species_nucLSU,
    nucSSU = species_nucSSU
  ),
  filename     = NULL,

  fill         = venn_cols,
  alpha        = 0.45,
  col          = "black",
  lwd          = 2,

  fontfamily       = "sans",
  cat.fontfamily   = "sans",
  main.fontfamily  = "sans",

  cex          = 1.8,
  fontface     = "bold",
  cat.cex      = 1.8,
  cat.fontface = "bold",
  cat.col      = venn_cols,

  cat.pos      = c(-20, 20, 180, 180),
  cat.dist     = c(0.22, 0.22, -0.12, -0.12),
  cat.just     = list(c(0.5,0.5), c(0.5,0.5), c(0.5,0.5), c(0.5,0.5)),
  main         = "",

  # >>> reduce white margins around the diagram <<<
  margin       = 0.03
)

pdf("Figures/Final Figures/Venn_markers_170mm.pdf",
    width  = 170/25.4,
    height = 140/25.4,
    family = "sans")

grid.newpage()

vp_width  <- 0.88

pushViewport(viewport(
  width  = vp_width,
  height = 0.9,               # slightly less than full height
  y      = 0.55,              # move the Venn up (0.5 = center)
  just   = c("center", "center")
))
grid.draw(venn.plot)
popViewport()

dev.off()


```

# Figure 4. Conservation Indices
## Prep
```{r}
library(ggplot2)
library(grid)

# Load data
superdict <- read.csv("SUPERDICT_28-08-2023(superdict).csv")

# Select and rename
cons_eval <- superdict %>% 
  select(Current.taxon.name, Cons.eval) %>%
  dplyr::rename(species = Current.taxon.name) %>%
  mutate(sequence = "test")

cons_eval <- fix_names(cons_eval, names)

# Clean up strings
cons_eval <- cons_eval %>%
  mutate(Cons.eval = trimws(Cons.eval),
         species = trimws(gsub("\\?", "", species))) %>%
  mutate(across(everything(), ~ na_if(.x, ""))) %>%
  filter(str_count(species, "\\S+") >= 2) %>%
  mutate(species = str_extract(species, "^\\S+\\s+\\S+"),
         Cons.eval = gsub("n/a", NA, Cons.eval)) %>%
  na.omit()

# Drop temp column
cons_eval <- cons_eval %>% select(-sequence)

# Add marker presence/absence
cons_eval <- cons_eval %>%
  mutate(
    ITS = ifelse(species %in% its_seq$species, 1, 0),
    mtssu = ifelse(species %in% mtssu_comb$species, 1, 0),
    nuclsu = ifelse(species %in% nuclsu_comb$species, 1, 0),
    nucssu = ifelse(species %in% nucSSU_seq$species, 1, 0)
  )

# Filter to relevant species
cons_eval <- cons_eval %>% 
  filter(species %in% names$new_name) %>%
  distinct()  # Remove duplicates

# Clean again
cons_eval$Cons.eval <- gsub("\\?", "", cons_eval$Cons.eval)

# Split IUCN codes into rows
valid_IUCN_codes <- c("LC", "NT", "VU", "EN", "CR", "DD", "NE", "EX")

# Split out IUCN codes, keep only valid ones
cons_eval_long <- cons_eval %>%
  mutate(ConsCodes = strsplit(Cons.eval, " ")) %>%
  unnest(ConsCodes) %>%
  filter(ConsCodes %in% valid_IUCN_codes)

# Count species before pivot
total_counts <- cons_eval_long %>%
  group_by(ConsCodes) %>%
  summarise(total_species = n_distinct(species), .groups = "drop")

# Pivot to long format for plotting
cons_eval_long <- cons_eval_long %>%
  pivot_longer(
    cols = c(ITS, mtssu, nuclsu, nucssu),
    names_to = "marker",
    values_to = "present"
  ) %>%
  group_by(ConsCodes, marker) %>%
  summarise(prop_with_seq = mean(present), .groups = "drop")

# Compute the maximum height of bars for each IUCN category
max_heights <- cons_eval_long %>%
  group_by(ConsCodes) %>%
  summarize(max_prop = max(prop_with_seq), .groups = "drop") %>%
  left_join(total_counts, by = "ConsCodes")

# Make sure ConsCodes has the correct order
cons_eval_long$ConsCodes <- factor(cons_eval_long$ConsCodes, levels = valid_IUCN_codes)
max_heights$ConsCodes <- factor(max_heights$ConsCodes, levels = valid_IUCN_codes)

```

## Graph
```{r}
library(ggplot2)
library(grid)

# Greyscale for the 4 markers
gray_colors <- c(
  "ITS"    = "grey0",
  "mtssu"  = "grey33",
  "nuclsu" = "grey66",
  "nucssu" = "grey100"
)

base_size_small <- 8

lichen_theme_small <- theme_classic(base_size = base_size_small) +
  theme(
    # outer frame only, same thickness on all sides
    panel.border      = element_rect(colour = "black", fill = NA, linewidth = 0.4),
    axis.line         = element_blank(),
    axis.line.x       = element_blank(),
    axis.line.y       = element_blank(),

    # ticks: thin + black, same both axes
    axis.ticks        = element_line(linewidth = 0.35, colour = "black"),
    axis.ticks.length = unit(3, "pt"),

    # text – all black; y-axis title centred on figure
    axis.title.x      = element_text(
      size   = 9,
      face   = "bold",
      margin = margin(t = 3),
      colour = "black"
    ),
    axis.title.y      = element_text(
      size   = 9,
      face   = "bold",
      margin = margin(r = 3),
      hjust  = 0.5,
      vjust  = 0.5,
      colour = "black"
    ),
    axis.text         = element_text(size = 8, colour = "black"),

    # LEGEND: top-right, with a box, slightly inset so it doesn't double the border
    legend.position      = c(0.98, 0.98),           # inside panel, near top-right
    legend.justification = c("right", "top"),
    legend.background    = element_rect(
      fill      = "white",
      colour    = "black",
      linewidth = 0.4
    ),
    legend.title         = element_blank(),
    legend.text          = element_text(size = 7, colour = "black"),
    legend.key           = element_rect(fill = NA, colour = NA),
    legend.key.width     = unit(8, "pt"),
    legend.key.height    = unit(8, "pt"),
    legend.margin        = margin(t = 2, r = 3, b = 2, l = 3),

    legend.box.margin    = margin(0, 0, 0, 0),
    legend.box.spacing   = unit(0, "pt"),

    plot.margin       = margin(4, 4, 4, 4)
  )

# ensure factor order for x axis
cons_eval_long$ConsCodes <- factor(cons_eval_long$ConsCodes,
                                   levels = valid_IUCN_codes)
max_heights$ConsCodes    <- factor(max_heights$ConsCodes,
                                   levels = valid_IUCN_codes)

# shared dodge so bars for the same category have small gaps
pd <- position_dodge(width = 0.8)  # a bit wider than the bar width

p_IUCN <- ggplot(cons_eval_long,
                 aes(x = ConsCodes, y = prop_with_seq, fill = marker)) +
  geom_col(
    position = pd,
    colour   = "black",
    linewidth = 0.25,
    width    = 0.6   # slightly narrower than dodge -> small gaps between bars
  ) +
  geom_text(
    data        = max_heights,
    aes(x = ConsCodes, y = max_prop + 0.02, label = total_species),
    inherit.aes = FALSE,
    vjust       = 0,
    fontface    = "bold",
    size        = 3,
    colour      = "black"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.15))
  ) +
  scale_fill_manual(
    values = gray_colors,
    labels = c("ITS" = "ITS",
               "mtssu" = "mtSSU",
               "nuclsu" = "nucLSU",
               "nucssu" = "nucSSU")
  ) +
  labs(
    x    = "IUCN category",
    y    = "Proportion of category with a sequenced species",
    fill = ""
  ) +
  lichen_theme_small +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

# quick check on screen
print(p_IUCN)

ggsave(
  filename = "Figures/Final Figures/Figure_IUCN_markers_84mm_v3.pdf",
  plot     = p_IUCN,
  device   = cairo_pdf,
  width    = 84/25.4,
  height   = 90/25.4,
  units    = "in",
  bg       = "white"
)


```


## Figure 5. Growth Form Taxon Dictionary
```{r}
bls_growth_forms <- read_xlsx("Taxonomy/bls-taxon-dictionary-2024-05-09_GROWTH FORMS.xlsx")

bls_growth_forms <- bls_growth_forms %>% select(`Current Taxon Name`, Group) %>%
  dplyr::rename(species = `Current Taxon Name`) %>% unique() %>% na.omit()

# Fix: remove weird group labels
bls_growth_forms <- bls_growth_forms %>%
  filter(!Group %in% c("A", "CBM", "CR", "Lepraria botryoides"))

# Fix: standardize short codes using exact matching
bls_growth_forms <- bls_growth_forms %>%
  mutate(Group = case_when(
    Group == "C"  ~ "Crustose",
    Group == "F"  ~ "Ambiguous",
    Group == "FO" ~ "Foliose",
    Group == "LF" ~ "Lichenicolous",
    Group == "B"   ~ "Fruticose",
    Group == "S"   ~ "Squamulose",
    TRUE ~ Group  # keep original if not matched
  ))


bls_growth_forms <- unique(bls_growth_forms)

# Function to get proportions per marker
get_growth_form_counts <- function(seq_df, marker_name, growth_forms_df) {
  seq_df %>%
    filter(species %in% growth_forms_df$species) %>%
    left_join(growth_forms_df, by = "species") %>%
    count(Group) %>%
    mutate(
      marker = marker_name,
      proportion = n / sum(n)
    )
}

bls_growth_forms <- bls_growth_forms %>% select(species, Group)

bls_growth_forms <- unique(bls_growth_forms)

growth_its <- its_seq %>% select(species) %>%
  filter(species %in% bls_growth_forms$species) %>%
  unique() %>%
  left_join(bls_growth_forms) %>%
  count(Group)%>%
  mutate(
    marker = "ITS",
    proportion = n / sum(n)
)

growth_nuclsu <- nuclsu_comb %>% select(species) %>%
  filter(species %in% bls_growth_forms$species) %>%
  unique() %>%
  left_join(bls_growth_forms) %>%
  count(Group)%>%
  mutate(
    marker = "nucLSU",
    proportion = n / sum(n)
)

growth_nucssu <- nucSSU_seq %>% select(species) %>%
  filter(species %in% bls_growth_forms$species) %>%
  unique() %>%
  left_join(bls_growth_forms) %>%
  count(Group)%>%
  mutate(
    marker = "nucSSU",
    proportion = n / sum(n)
)

growth_mtssu <- mtssu_comb %>% select(species) %>%
  filter(species %in% bls_growth_forms$species) %>%
  unique() %>%
  left_join(bls_growth_forms) %>%
  count(Group) %>%
  mutate(
    marker = "mtSSU",
    proportion = n / sum(n)
)


# Add All as a reference bar using full dataset
growth_all_total <- bls_growth_forms %>%
  count(Group) %>%
  mutate(
    marker = "All",
    proportion = n / sum(n)
  )

# Combine all
growth_all <- bind_rows(growth_its, growth_nuclsu, growth_nucssu, growth_mtssu, growth_all_total)

# Factor levels for stacking largest to smallest within each bar
# Get average proportions across markers for sorting
growth_order <- growth_all %>%
  group_by(Group) %>%
  summarise(avg = mean(proportion)) %>%
  arrange(desc(avg)) %>%
  pull(Group)

# Set custom stacking order for growth forms (bottom to top in barplot)
growth_all$Group <- factor(growth_all$Group,
                           levels = c("Squamulose","Ambiguous", "Fruticose", "Foliose", "Lichenicolous", "Crustose"))
library(ggplot2)
library(scales)
library(grid)

# Greyscale for the 6 growth forms
growth_greys <- c(
  "Squamulose"    = "grey90",
  "Ambiguous"     = "grey72",
  "Fruticose"     = "grey54",
  "Foliose"       = "grey36",
  "Lichenicolous" = "grey18",
  "Crustose"      = "grey0"
)

legend_labels <- c(
  "Squamulose"    = "Squamulose",
  "Ambiguous"     = "Ambiguous",
  "Fruticose"     = "Fruticose",
  "Foliose"       = "Foliose",
  "Lichenicolous" = "Lichenicolous",
  "Crustose"      = "Crustose"
)

# ---------- Theme to match Figures 2 & 4 ----------
base_size_small <- 8
lichen_theme_small2 <- theme_classic(base_size = base_size_small) +
  theme(
    # outer frame only, thinner and uniform
    panel.border      = element_rect(color = "black", fill = NA, linewidth = 0.4),
    axis.line         = element_blank(),

    # ticks: thin + black, same both axes
    axis.ticks        = element_line(linewidth = 0.35, colour = "black"),
    axis.ticks.length = unit(3, "pt"),

    # tick labels: small, all black
    axis.text         = element_text(size = 7, colour = "black"),

    # axis titles: all black, bold
    axis.title.x = element_text(
      size   = 8,
      face   = "bold",
      margin = margin(t = 4),
      colour = "black"
    ),
    axis.title.y = element_text(
      size   = 8,
      face   = "bold",
      hjust  = 0.5,
      vjust  = 0.5,
      margin = margin(r = 0),   # <<< closer to plot (was 3)
      colour = "black"
    ),

    # legend on right, a bit closer to the plot
    legend.position    = "right",
    legend.title       = element_text(size = 8, face = "bold", colour = "black"),
    legend.text        = element_text(size = 7, colour = "black"),
    legend.key.width   = unit(8, "pt"),
    legend.key.height  = unit(8, "pt"),
    legend.spacing.y   = unit(2, "pt"),          # small vertical separation
    legend.box.margin  = margin(0, 0, 0, 0),     # <<< tighter to plot (was 0,2,0,2)

    # overall plot margins: pull right side in slightly
    plot.margin        = margin(4, 0.5, 4, 1)    # top, right, bottom, left
  )


p_growth <- ggplot(
  growth_all,
  aes(x = marker,
      y = proportion,
      fill = Group)
) +
  geom_col(
    position = "stack",
    colour   = "white",
    linewidth = 0.2,   # thin internal segment/boundary lines
    width    = 0.75
  ) +
  scale_fill_manual(
    values = growth_greys,
    labels = legend_labels,
    name   = "Growth form"
  ) +
  scale_y_continuous(
    labels = percent_format(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Marker",
    y = "Species proportion"
  ) +
  lichen_theme_small2

# ---------- Save as 84 mm PDF ----------
ggsave(
  filename = "Figures/Final Figures/Figure_growth_forms_84mm.pdf",
  plot     = p_growth,
  device   = cairo_pdf,
  width    = 84/25.4,       # convert mm → inches
  height   = 95/25.4,
  units    = "in",
  bg       = "white"
)

```


# Priority Taxa list
## Family
```{r}
# Abundace
bls_sp_counts <- read_excel("FW_ Species counts by country/BLS species counts by VC Sept 2024.xlsx")

# Fix and clean
colnames(bls_sp_counts) <- c("BLS number", "Taxon", "Vice county", bls_sp_counts[1, 4:ncol(bls_sp_counts)])
bls_sp_counts <- bls_sp_counts[-1,]

names(bls_sp_counts) <- make.unique(names(bls_sp_counts))

bls_sp_counts <- bls_sp_counts %>%
  dplyr::select(-`Vice county`, -`BLS number`)

# Make every column numeric
sp_counts <- rowSums(apply(bls_sp_counts[, 2:ncol(bls_sp_counts)], 2, as.numeric), na.rm = T)

bls_sp_counts_final <- data.frame(species = bls_sp_counts[,1], counts = sp_counts)

# Totals
bls_sp_counts_final <- bls_sp_counts_final[c(-2550, -2551, -2552), ]
bls_sp_counts_final <- na.omit(bls_sp_counts_final)

# Just take species
bls_sp_counts_final$species <- sapply(strsplit(bls_sp_counts_final$Taxon, "\\s+"), function(words) {
  paste(words[1:min(2, length(words))], collapse = " ")
})

# group by and sum by Taxon
bls_sp_counts_final <- bls_sp_counts_final %>% group_by(species) %>%
  summarize(counts = sum(counts))

# Remove rows where counts is 0
bls_sp_counts_final <- bls_sp_counts_final %>%
  filter(counts > 0)

# Tax data
tax_ITS <- tax_df %>% filter(species %in% its_seq$species) %>% left_join(bls_sp_counts_final)
tax_ssu <- tax_df %>% filter(species %in% mtssu_comb$species) %>% left_join(bls_sp_counts_final)
tax_lsu <- tax_df %>% filter(species %in% nuclsu_comb$species)%>% left_join(bls_sp_counts_final)
tax_all <- tax_df %>% filter(species %in% names$new_name)%>% left_join(bls_sp_counts_final)
tax_all$counts <- tax_all$counts %>% replace_na(0)
family_abund <- tax_all %>% group_by(family) %>%  summarize(abund = sum(counts))

count_ITS  <- sort(table(tax_ITS$class))
count_ssu  <- sort(table(tax_ssu$class))
count_lsu  <- sort(table(tax_lsu$class))
count_all  <- sort(table(tax_all$class))

# Get unique categories (classes) from all datasets
categories <- unique(c(names(count_ITS), names(count_ssu), names(count_lsu), names(count_all)))

# Summarize each dataset
tax_ITS_counts <- tax_ITS %>% 
  group_by(family) %>% 
  summarize(n_ITS = n(), .groups = "drop")

tax_ssu_counts <- tax_ssu %>% 
  group_by(family) %>% 
  summarize(n_ssu = n(), .groups = "drop")

tax_lsu_counts <- tax_lsu %>% 
  group_by(family) %>% 
  summarize(n_lsu = n(), .groups = "drop")

tax_all_counts <- tax_all %>% 
  group_by(family) %>% 
  summarize(n_all = n(), .groups = "drop")

# Full join them all, so families in ANY dataset are kept
family_counts <- tax_ITS_counts %>%
  full_join(tax_ssu_counts, by = "family") %>%
  full_join(tax_lsu_counts, by = "family") %>%
  full_join(tax_all_counts, by = "family") %>%
  full_join(family_abund, by = "family")

# Replace NA with 0 in numeric count columns
family_counts <- tax_ITS_counts %>%
  full_join(tax_ssu_counts, by = "family") %>%
  full_join(tax_lsu_counts, by = "family") %>%
  full_join(tax_all_counts, by = "family") %>%
  full_join(family_abund, by = "family") %>%
  mutate(
    n_ITS = replace_na(n_ITS, 0),
    n_ssu = replace_na(n_ssu, 0),
    n_lsu = replace_na(n_lsu, 0),
    n_all = replace_na(n_all, 0),
    abund = replace_na(abund, 0),
    n_sum = n_ITS + n_ssu + n_lsu
  ) %>%
  filter(!is.na(family)) %>%
  filter(!str_detect(family, "incertae_sedis"))

families_without_sequences <- family_counts %>%
  filter(n_sum == 0)

family_prior <- families_without_sequences %>%
  select(family) %>%
  left_join(tax_df, by = "family") %>%
  left_join(bls_sp_counts_final, by = "species") %>%
  mutate(counts = replace_na(counts, 0)) %>%
  group_by(family) %>%
  slice_max(counts, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(family, species, counts)

#write.csv(family_prior, "C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Updater/March_2026/Priority_lists/family_priority_list_Mar_6_26.csv", row.names = F)

```

## Conservation 
```{r}

# Load data
superdict <- read.csv("SUPERDICT_28-08-2023(superdict).csv")

# Select and rename
cons_eval <- superdict %>% 
  select(Current.taxon.name, Cons.eval) %>%
  dplyr::rename(species = Current.taxon.name) %>%
  mutate(sequence = "test")

# Optional: if you have a custom fix_names() function, uncomment this line
cons_eval <- fix_names(cons_eval, names)

# Clean up strings
cons_eval <- cons_eval %>%
  mutate(Cons.eval = trimws(Cons.eval),
         species = trimws(gsub("\\?", "", species))) %>%
  mutate(across(everything(), ~ na_if(.x, ""))) %>%
  filter(str_count(species, "\\S+") >= 2) %>%
  mutate(species = str_extract(species, "^\\S+\\s+\\S+"),
         Cons.eval = gsub("n/a", NA, Cons.eval)) %>%
  na.omit()

# Drop temp column
cons_eval <- cons_eval %>% select(-sequence)

# Add marker presence/absence
cons_eval <- cons_eval %>%
  mutate(
    ITS = ifelse(species %in% its_seq$species, 1, 0),
    mtssu = ifelse(species %in% mtssu_comb$species, 1, 0),
    nuclsu = ifelse(species %in% nuclsu_comb$species, 1, 0),
    nucssu = ifelse(species %in% nucSSU_seq$species, 1, 0)
  )

# Filter to relevant species
cons_eval <- cons_eval %>% 
  filter(species %in% names$new_name) %>%
  distinct()  # Remove duplicates

# Clean agan
cons_eval$Cons.eval <- gsub("\\?", "", cons_eval$Cons.eval)

# --- PLOT PREP ---
# STEP A: Split IUCN codes into rows
valid_IUCN_codes <- c("LC", "NT", "VU", "EN", "CR", "DD", "NE", "EX")

# Split out IUCN codes, keep only valid ones
cons_eval_long <- cons_eval %>%
  mutate(ConsCodes = strsplit(Cons.eval, " ")) %>%
  unnest(ConsCodes) %>%
  filter(ConsCodes %in% valid_IUCN_codes)

sp_seq <- cons_eval_long %>% filter(ITS == 0)

sp_seq <- sp_seq %>% filter(ConsCodes == "CR" | ConsCodes == "EN" | ConsCodes == "DD")

sp_seq <- sp_seq %>% left_join(bls_sp_counts_final)

sp_seq <- sp_seq %>% select(species, ConsCodes, counts)

sp_seq <- sp_seq %>% mutate(counts = replace_na(counts, 0))

# write.csv(sp_seq, "C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Updater/March_2026/Priority_lists/conservation_priority_list_Mar_6_26.csv", row.names = F)

```

## Most abundant
```{r}

missing_its <- names %>% filter(!new_name %in% its_seq$species) %>% dplyr::rename(species = new_name) %>% select(-old_name)

missing_its <- missing_its %>% left_join(bls_sp_counts_final)

missing_its <- unique(missing_its)

top_100_missing_its <- missing_its %>%
  slice_max(order_by = counts, n = 100)

#write.csv(top_100_missing_its, "C:/Users/dylan/OneDrive/Desktop/PhD/gap_analysis/Updater/March_2026/Priority_lists/most_abundant_priority_list_Mar_6_26.csv", row.names = F)
```


























